---
title: "Phase II - District Characteristics by Change Group"
author: "Janette Avelar"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: flatly
    highlight: tango
    code_folding: hide
  word_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # hides code
  message = FALSE, # hides messages
  warning = FALSE,  # hides warnings
  fig.width = 10, fig.height = 9
)
library(rio)
library(tidyverse)
library(here)
library(DT)
library(kableExtra)

dat <- import('/Users/janetteavelar/Library/CloudStorage/GoogleDrive-javelar3@asu.edu/.shortcut-targets-by-id/1NdfrCmhAW0NpUV47ruVeKThD8M3arhRq/DATA: Barr 3.0/Survey/Analysis/Current Data Files/MergedSurvey_251215.dta')
source(here("functions/change_groups_functions.R"))
```

# Survey Task

In this analysis, we are interested in the guiding question: **How does context relate to change groups?**  

There are 4 change groups, defined as:  
- 0 changes underway  
- 1-2 changes underway  
- 3-5 changes underway  
- 6-9 changes underway  

This file looks at district characteristics, comparing the mean and distribution of state, district size (enrollment), district size (number of schools), serve high school, enrollment change, racial composition (% White), % low inc, % EL % SWD, student outcomes by change group.  

**Note**: *There are 8 districts that do not fall into a change group and I'm not totally sure why -- I checked their responses for Q4-6, and 2 of the districts had responded to all 3, while some seem due to missingness. Would be good to talk about who these districts are and why they're missing.*  

```{r}
# overview
dat %>% 
  filter(is.na(change_group)) %>% 
  select(qcode, partial, state, total_domains_working,change_group) %>% 
  kable()

# q6
tmp <- dat %>%
  filter(is.na(change_group)) %>%
  select(qcode, starts_with("q6"))

highlight <- which(!is.na(tmp$q6_2))

tmp %>%
  kable() %>%
  kable_styling() %>%
  row_spec(highlight, color = "black", background = "yellow")
rm(tmp, highlight)
```


```{r}
# Clean up data
clean <- dat %>% 
  haven::zap_label() %>% 
  mutate(grades_served = case_when(
    grades_served == 1 ~ "K-6",
    grades_served == 2 ~ "K-8",
    grades_served == 3 ~ "K-12",
    grades_served == 4 ~ "6-8",
    grades_served == 5 ~ "7-12",
    grades_served == 6 ~ "9-12"
  ),
  district_type = case_when(
    district_type == 1 ~ "Traditional",
    district_type == 2 ~ "Regional",
    district_type == 3 ~ "Regional CTE",
    district_type == 4 ~ "Charter",
    district_type == 5 ~ "State agency"
  ),
  urbanicity = case_when(
    urbanicity == 1 ~ "Urban",
    urbanicity == 2 ~ "Suburban",
    urbanicity == 3 ~ "Rural"
  ),
  ma_account_code = case_when(
    ma_account_code == 1 ~ "Insufficient data",
    ma_account_code == 2 ~ "Not requiring assistance",
    ma_account_code == 3 ~ "Requiring assistance or intervention"
  ),
  change_group = case_when(
    change_group == 1 ~ "0 domains",
    change_group == 2 ~ "1-2 domains",
    change_group == 3 ~ "3-5 domains",
    change_group == 4 ~ "6-9 domains"
  ),
  serve_hs = case_when(
    serve_hs == 0 ~ "Does not serve high school",
    serve_hs == 1 ~ "Serves high school"
  )) %>% 
  filter(!is.na(change_group))
```

# A1 By state

There are 6 states in our data:  

```{r}
table(clean$state) %>% kable()
```

The percentages in this table represent the distribution within a given state, e.g., *"In Connecticut, 18% of districts were not undergoing change (0 domains) while 6% had high change (6-9 domains)."*  

The pattern of distributions is fairly even, but there are some differences to note:  
- Maine and Rhode Island both are skewed toward less, or more focused, change, with the majority (42% & 47% respectively) of districts in the state focusing on 1 or 2 changes.  
- New Hampshire had the highest proportion of districts not focusing on change at all (0 domains - 20%) and the smallest proportion of districts undergoing a lot of change (6-9 domains - 2%).  
- Vermont had the smallest proportion of districts undergoing no change (0 domains - 8%).  

```{r}
# generate counts - distribution of change groups within state (rowwise)
state <- clean %>% 
  group_by(change_group) %>% 
  count(state) %>% 
  ungroup() %>% 
  group_by(state) %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  # Note: rounding to 2 decimals was equivalent when testing pct overall or pct within group
  mutate(change_group = ifelse(is.na(change_group), "Missing", change_group),
         pct = round(n/total, 2),
         state = case_when(
           state == "CT" ~ "Connecticut",
           state == "MA" ~ "Massachusetts",
           state == "ME" ~ "Maine",
           state == "NH" ~ "New Hampshire",
           state == "RI" ~ "Rhode Island",
           state == "VT" ~ "Vermont"
         )) %>% 
  select(change_group, state, pct) %>% 
  pivot_wider(names_from = change_group,
              values_from = pct)

# table
datatable(state,
          caption = "Table A.1 Change groups by state",
          options = list(searchHighlight = TRUE))
```

```{r}
# viz 1
state %>% 
  pivot_longer(cols = !state,
               names_to = "change_group",
               values_to = "pct") %>% 
  ggplot(., aes(pct, state, fill = change_group)) +
  geom_col(position = position_stack(reverse = TRUE), aes(group = change_group)) +
  theme_bw() +
  labs(title = "Distribution of change groups by state",
       x = "Percent of districts",
       y = "",
       fill = "Change group") +
  scale_fill_manual(values = c("#f0b8b8", "#e67f83", "#aecdc2", "#1F916F")) +
  geom_text(aes(x = ave(pct, state, FUN = cumsum) - pct/2,
                label = ifelse(!is.na(pct) & pct > 0, scales::percent(pct, accuracy = 1), "")),
            position = "identity",
            size = 3.5,
            show.legend = FALSE) 
```

When we break out the distribution of states within change group, we see that high change (3-5 domains or 6-9 domains) is concentrated in Massachusetts (37% and 42% respectively). 

```{r results='asis'}
# generate counts - distribution of states within change group (colwise)
state_alt <- cat_table(clean, change_group, state, "Table A.1 Distribution of states within change group")
```

```{r}
# viz 2
clean %>% 
  group_by(change_group) %>% 
  count(state) %>% 
  mutate(pct = n/sum(n)) %>% 
  ungroup() %>% 
  ggplot(., aes(state, pct)) +
  geom_col(fill = "#1F916F") +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 1)) +
  scale_x_discrete(drop = FALSE) +
  theme_bw() +
  labs(title = "Distribution of states by change group",
       x = "Percent",
       y = "") +
  geom_text(
      aes(label = scales::percent(pct, accuracy = 1)),
      vjust = -1,
      size = 3,
      show.legend = FALSE
    ) +
  facet_wrap(~change_group) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
``` 

We also want to know if these means are statistically significant from one another. Given we are accounting for state, I ran a chi-square test. Although the share of districts in each change group varies across states, there is no evidence the distribution differs significantly by change group across states.

```{r}
# chi square test
tab <- table(clean$state, clean$change_group)
chisq <- chisq.test(tab)
chisq
chisq$stdres
```


```{r include = FALSE, eval = FALSE}
# save table
write.csv(state, "output/tables/table_a1.csv", row.names = FALSE)
rm(state, state_alt, tab, chisq)
```

# A2 By district size

## Defined as enrollment

We have enrollment information for the 2024-25 school across 304 districts. The table below includes descriptive statistics across the broader sample:

```{r}
# overall distribution
psych::describe(clean$enrollment_2025) %>% 
  select(n, mean, sd, min, max) %>% 
  kable()
```

Table A2.1 displays descriptive statistics broken out by the number of domains a given district is working on.

```{r}
enroll <- cont_table(clean, change_group, enrollment_2025, c(mean, sd), "A2.1 Enrollment by Change Group")
```

```{r}
# viz
cont_boxplot(clean, change_group, enrollment_2025, 
             title = "Distribution of enrollment in 2025 by change group", 
             x_lab = "", 
             y_lab = "Enrollment")
```

```{r}
# alt viz
# SHOULD I ADD PERCENTAGES???
# If yes, need to think about denominator - do we want it to be among domain group, or among enrollment band?
clean %>% 
  mutate(enroll_group = case_when(
    enrollment_2025 < 500 ~ "< 500",
    enrollment_2025 >= 500 & enrollment_2025 < 1500 ~ "500-1,499",
    enrollment_2025 >= 1500 & enrollment_2025 < 3000 ~ "1,500-2,999",
    enrollment_2025 >= 3000 & enrollment_2025 < 7500 ~ "3,000-7,499",
    #enrollment_2025 >= 5000 & enrollment_2025 < 7500 ~ "5,000-7,500",
    enrollment_2025 >= 7500 ~ "> 7,500"
  )) %>% 
  group_by(change_group) %>% 
  count(enroll_group) %>% 
  filter(!is.na(enroll_group)) %>% 
  filter(!is.na(change_group)) %>% 
  mutate(enroll_group = factor(enroll_group, levels = c("> 7,500", "3,000-7,499", "1,500-2,999", "500-1,499", "< 500"))) %>% 
  ggplot(., aes(n, enroll_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0),
                     limits = c(0, 50)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 1.5),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of enrollment size by change group",
       x = "Number of districts",
       y = "Number of students")
```

There is not strong statistical evidence that enrollment size differs by change group, although the overall test is marginal (p ≈ .055).

```{r}
summary(lm(enrollment_2025 ~ change_group, data = clean))
```


```{r, include = FALSE, eval = FALSE}
# save table
write.csv(enroll, "output/tables/table_a2.1.csv", row.names = FALSE)
rm(enroll)
```


## Defined as number of schools

The table below includes descriptive statistics for the number of schools and number of high schools districts have:

```{r}
# overall descriptives
subset <- clean %>% 
  select(num_schools, num_high_schools)
psych::describe(subset) %>% 
  select(n, mean, sd, min, max) %>% 
  kable()
rm(subset)
```

```{r}
# generating descriptives by group
n_school <- clean %>% 
  group_by(change_group) %>% 
  summarize(mean_high_schools = mean(num_high_schools, na.rm = TRUE),
            sd_high_schools = sd(num_high_schools, na.rm = TRUE),
            min_high_schools = min(num_high_schools, na.rm = TRUE),
            max_high_schools = max(num_high_schools, na.rm = TRUE),
            mean_all_schools = mean(num_schools, na.rm = TRUE),
            sd_all_schools = sd(num_schools, na.rm = TRUE),
            min_all_schools = min(num_schools, na.rm = TRUE),
            max_all_schools = max(num_schools, na.rm = TRUE)) %>% 
  pivot_longer(cols = -change_group,
               names_to = c("stat", "group"),
               names_pattern = "(mean|sd|min|max)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat,
              values_from = value) %>%
  mutate(group = recode(group,
                        high_schools = "High school",
                        all_schools  = "All schools"),
         change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  mutate(across(c(mean, sd), ~round(., 2)))

# table
datatable(n_school,
          caption = "A2.2 Distribution of number of total schools and high schools by change group")
```

```{r}
# viz
clean %>% 
  select(change_group, num_schools, num_high_schools) %>% 
  pivot_longer(cols = c("num_schools", "num_high_schools"),
               names_to = "group",
               values_to = "n_school",
               names_prefix = "num_") %>% 
  mutate(group = ifelse(group == "schools", "All schools", "High schools"),
         change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  ggplot(., aes(change_group, n_school)) +
  geom_boxplot() +
  facet_wrap(~group) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       subtitle = str_wrap("These boxplots represent the median (bold line), 25th percentile (bottom of box), 75th percentile (top of box), and outliers (dots).", 100),
       x = "",
       y = "Number of schools")
```

```{r}
# alt viz
clean %>% 
  group_by(change_group) %>% 
  #count(num_schools) %>% 
  mutate(num_schools_group = case_when(
    num_schools < 5 ~ "< 5",
    num_schools >= 5 & num_schools < 10 ~ "5-9",
    num_schools >= 10 & num_schools < 15 ~ "10-14",
    num_schools >= 15 & num_schools < 20 ~ "15-19",
    num_schools >= 20 & num_schools < 25 ~ "20-24",
    num_schools >= 25 ~ "25+"
  ),
  num_schools_group = factor(num_schools_group, levels = c("25+", "20-24", "15-19", "10-14", "5-9", "< 5"))) %>% 
  count(num_schools_group) %>% 
  filter(!is.na(change_group)) %>% 
  ggplot(., aes(n, num_schools_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 2),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       x = "Number of districts",
       y = "Number of schools in district")
```

There is not strong statistical evidence that enrollment size differs by the number of schools in the district (p ≈ .06).

```{r}
summary(lm(num_schools ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
write.csv(n_school, "output/tables/table_a2.2.csv", row.names = FALSE)
```


# A3 By serving high school

```{r}
# alt viz
clean %>% 
  group_by(change_group) %>% 
  #count(num_high_schools) 
  mutate(num_schools_group = case_when(
    num_high_schools == 0 ~ "0",
    num_high_schools == 1 ~ "1",
    num_high_schools == 2 ~ "2",
    num_high_schools == 3 ~ "3",
    num_high_schools >= 4 ~ "4+"
  ),
  num_schools_group = factor(num_schools_group, levels = c("4+", "3", "2", "1", "0"))) %>% 
  count(num_schools_group) %>% 
  filter(!is.na(change_group)) %>% 
  ggplot(., aes(n, num_schools_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 2),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       x = "Number of districts",
       y = "Number of high schools in district")
```

There is not strong statistical evidence that enrollment size differs by the number of high schools in the district (p ≈ .18).

```{r}
summary(lm(num_high_schools ~ change_group, data = clean))
```

```{r}
# clean up
rm(n_school)
```


## By enrollment change

We have 6-year enrollment data for 296 districts (11 missing) indicating enrollment change over that time. Enrollment is trending downwards in our data, with the average district dropping 3 percentage points (check interpretation) between 2019 and 2025. The majority of districts (74%) had a value below 0, indicating a loss in enrollment.

```{r}
# overall descriptives
psych::describe(clean$enrollment_change_6yr_pct)
# 26% gaining, 74% losing; 11 districts held missing values
clean %>% 
  mutate(group = ifelse(enrollment_change_6yr_pct <= 0, "Loss", "Gain")) %>% 
  #count(group)
  ggplot(., aes(enrollment_change_6yr_pct)) +
  geom_histogram(fill = "#1F916F") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  theme_bw() +
  labs(title = "Distribution of 6-year enrollment change across our sample",
       subtitle = "Red dashed line indicates zero change.",
       x = "6-year enrollment change",
       y = "Number of districts")
```

When we break out enrollment change by change group, we don't see any obvious differences in the pattern of loss. Districts with a higher degree of change (6-9 domains) are experiencing a slightly higher rate of loss compared to other groups (6%).

```{r}
# generating descriptives by group
enroll_change <- cont_table(clean, change_group, enrollment_change_6yr_pct, c(mean, sd, min, max), "A2.3 Distribution of enrollment change by change group")
```

```{r}
# first viz
clean %>% 
  filter(!is.na(change_group)) %>% 
  mutate(group = ifelse(enrollment_change_6yr_pct <= 0, "Loss", "Gain")) %>% 
  #count(group)
  ggplot(., aes(enrollment_change_6yr_pct)) +
  geom_histogram(fill = "#1F916F") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  theme_bw() +
  labs(title = "Distribution of 6-year enrollment change among change groups",
       subtitle = "Red dashed line indicates zero change.",
       x = "6-year enrollment change",
       y = "Number of districts") +
  facet_wrap(~change_group)
```

```{r}
# second viz
clean %>% 
  filter(!is.na(change_group)) %>% 
  mutate(group = ifelse(enrollment_change_6yr_pct <= 0, "Loss", "Gain")) %>% 
  group_by(change_group) %>% 
  filter(!is.na(group)) %>% 
  count(group) %>% 
  mutate(pct = n/sum(n),
         pct = ifelse(group == "Loss", pct*-1, pct)) %>% 
  ggplot(., aes(pct, change_group, fill = group)) +
  geom_col() +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  scale_fill_manual(values = c("#1F916F", "#e67f83")) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Distribution of districts with enrollment gains or loss between 2019 and 2025 across change groups",
       subtitle = "Red dashed line indicates zero change.",
       x = "Percent of districts",
       y = "",
       fill = "Enrollment trend") +
  geom_text(
  aes(label = scales::percent(abs(pct), accuracy = 1),
      hjust = ifelse(group == "Loss", 1.05, -0.05)),
  position = position_stack(vjust = .5),
  size = 3.5,
  show.legend = FALSE) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

There is not strong statistical evidence that the 6-year change in enrollment differs by the number of schools in the district (p ≈ .4).

```{r}
summary(lm(enrollment_change_6yr_pct ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
write.csv(enroll_change, "output/tables/table_a2.3.csv", row.names = FALSE)
rm(enroll_change)
```

# A3 By grades served

We have data for grades served for 299 districts (0 missing). The categories are broken into:  
- K-6  
- K-8  
- K-12  
- 6-8  
- 7-12  
- 9-12  

Counts by grade band are displayed in the table below:

```{r}
#overall descriptives
table(clean$grades_served)
```

Unsurprisingly, once we break out by change group we see that the majority of districts are falling into the K-12 grade band by change group. This is overwhelmingly the case for those districts implementing a lot of change (3-5 or 6-9 domains). Among districts focusing on just 1 or 2 domains, 1 in 5 is a district that serves only K-8 grades. Among districts that are not currently focusing on any domains, a quarter serve only K-6 or K-8 grades.  

This could indicate that districts serving high school grades, and particularly full K-12 grades, tend to take on more change work while those serving primary and middle school grades focus on far less, if any, change work.  

```{r}
#generate table & display
grades <- cat_table(clean, change_group, grades_served, "A.3 Grades served in district by change group")
```

The chart below breaks out percentages differently than the table above. Here, the distribution of change groups is broken out by grades served (rather than looking at grades served broken out by change group).  

```{r}
# viz
cat_stacked_bar(grades, grades_served, 
                title = "Distribution of change groups by the grades served in the district",
                x_lab = "Percent of districts", y_lab = "")
```

According to the chi-square test, the change groups do differ meaningfully based on the grades served in the district (p < .05). My initial thoughts are that this is due to cell sizes with 0 values as well as overlap in the grade bands. For future research, I'd suggest using the "serves high school grades" variable only.  

```{r}
# chi square test
tab <- table(clean$grades_served, clean$change_group)
chisq <- chisq.test(tab)
chisq
chisq$stdres
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(grades, "output/tables/table_a3.csv", row.names = FALSE)
rm(grades, tab, chisq)
```


# A4 By district type

We have district type data for all districts, and counts by type are in the table below:

```{r}
#overall
table(clean$district_type)
```

When we break this out by group, traditional schools make up the largest proportion of each change group. Other notable patterns include:
- Nearly 15% of districts working on 1-2 domains are regional districts, as are nearly 20% of those working on 1-2 domains  
- Charters are more highly represented among districts with no or focused efforts: 12% working on 0 domains, and 17% working on 1-2 domains

```{r}
# pull values
type <- cat_table(clean, change_group, district_type, "A.4 District type by change group")
```

The chart below breaks out percentages differently than the table above. Here, the distribution of change groups is broken out by school type (rather than looking at school type broken out by change group).  

```{r}
# viz
cat_stacked_bar(type, district_type, 
                title = "Distribution of change groups by district type", 
                x_lab = "Percent of districts", 
                y_lab = "")
```

There is not evidence that change groups differ meaningfully by district type according to the chi square results below (p = .101). 

```{r}
# chi square test
tab <- table(clean$district_type, clean$change_group)
chisq <- chisq.test(tab)
chisq
chisq$stdres
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(type, "output/tables/table_a4.csv", row.names = FALSE)
rm(type, tab, chisq)
```

# A5 By urbanicity

We have urbanicity data for all districts, and counts by type are in the table below:

```{r}
#overall
table(clean$urbanicity)
```

When we break this out by group, rural areas make up the highest proportion for all groups except for those with 6-9 domains. Among those with 6-9 domains, nearly half are suburban districts.

```{r}
# pull values
urbanicity <- cat_table(clean, change_group, urbanicity, "A.5 Urbanicity by change group")
```

The chart below breaks out percentages differently than the table above. Here, the distribution of change groups is broken out by urbanicity (rather than looking at urbanicity broken out by change group).  

```{r}
# viz
cat_stacked_bar(urbanicity, urbanicity,
                title = "Distribution of change groups by urbanicity",
                x_lab = "Percent of districts",
                y_lab = "")
```

There is not evidence that change groups differ meaningfully by urbanicity according to the chi square results below (p = .28). 

```{r}
# chi square test
tab <- table(clean$urbanicity, clean$change_group)
chisq <- chisq.test(tab)
chisq
chisq$stdres
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(urbanicity, "output/tables/table_a5.csv", row.names = FALSE)
rm(urbanicity, tab, chisq)
```


# A6 By racial composition

We have racial demographics data for about 304 districts (3 missing overall, 4 missing American Indian and Alaska Native data), including:  
- % White students  
- % Black students  
- % Latine/Hispanic students  
- % Asian or Pacific Islander students  
- % American Indian or Alaska Native  
- % Multiracial  

The overall descriptives for our sample are in the table below:

```{r}
# overall descriptives
clean %>% 
  select(pct_white, pct_black, pct_hispanic, pct_aapi, pct_aian, pct_multiracial) %>% 
  psych::describe()
```

When we break out descriptives by change group we see that districts undergoing more change (6-9 domains) tend to have higher populations of BIPOC students and combined Black/Hispanic student populations.  

```{r}
race <- clean %>% 
  mutate(bipoc = 1-pct_white,
         blackhispanic = pct_black+pct_hispanic) %>% 
  group_by(change_group) %>% 
  summarize(mean_white = mean(pct_white, na.rm = TRUE),
            sd_white = sd(pct_white, na.rm = TRUE),
            mean_black = mean(pct_black, na.rm = TRUE),
            sd_black = sd(pct_black, na.rm = TRUE),
            mean_hispanic = mean(pct_hispanic, na.rm = TRUE),
            sd_hispanic = sd(pct_hispanic, na.rm = TRUE),
            mean_aapi = mean(pct_aapi, na.rm = TRUE),
            sd_aapi = sd(pct_aapi, na.rm = TRUE),
            mean_aian = mean(pct_aian, na.rm = TRUE),
            sd_aian = sd(pct_aian, na.rm = TRUE),
            mean_multiracial = mean(pct_multiracial, na.rm = TRUE),
            sd_multiracial = sd(pct_multiracial, na.rm = TRUE),
            mean_bipoc = mean(bipoc, na.rm = TRUE),
            sd_bipoc = sd(bipoc, na.rm = TRUE),
            mean_blackhispanic = mean(blackhispanic, na.rm = TRUE),
            sd_blackhispanic = sd(blackhispanic, na.rm = TRUE)) %>% 
  mutate(change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  pivot_longer(cols = !change_group,
               names_to = c("stat", "race"),
               names_pattern = "(mean|sd)_(.*)",
               values_to = "value") %>% 
  pivot_wider(names_from = stat,
              values_from = value) %>% 
  pivot_wider(names_from = change_group,
              values_from = c(mean, sd)) %>% 
  mutate(race = case_when(
    race == "white" ~ "White",
    race == "black" ~ "Black",
    race == "hispanic" ~ "Hispanic",
    race == "aapi" ~ "Asian or Pacific Islander",
    race == "aian" ~ "American Indian or Alaska Native",
    race == "multiracial" ~ "Multiracial",
    race == "bipoc" ~ "All BIPOC students",
    race == "blackhispanic" ~ "Combined Black + Hispanic"
  )) %>% 
  mutate(across(!race, ~round(., 2))) %>% 
  select(race, `mean_0 domains`, `sd_0 domains`, `mean_1-2 domains`, `sd_1-2 domains`, `mean_3-5 domains`, `sd_3-5 domains`, `mean_6-9 domains`, `sd_6-9 domains`) 
# display
sketch = htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(colspan = 1, ""),
      th(colspan = 2, '0 domains'),
      th(colspan = 2, '1-2 domains'),
      th(colspan = 2, '3-5 domains'),
      th(colspan = 2, "6-9 domains")
    ),
    tr(
      lapply(c("Group", rep(c('Mean', 'SD'), 4)), th)
    )
  )
))
datatable(race,
          caption = "A6.1 District racial demographics by change group",
          colnames = c("", rep(c("Mean", "SD"), 4)),
          container = sketch,
          rownames = FALSE)
```

```{r, include = FALSE, eval = FALSE}
write.csv(race, "output/tables/table_a6.1.csv", row.names = FALSE)
rm(race, sketch)
```

```{r}
# viz 1 - boxplots
clean %>% 
  select(change_group, pct_white, pct_black, pct_hispanic, pct_aapi, pct_aian, pct_multiracial) %>% 
  mutate(pct_bipoc = 1-pct_white,
         pct_blackhispanic = pct_black+pct_hispanic) %>% 
  pivot_longer(cols = !change_group,
               names_to = "race",
               values_to = "pct",
               names_prefix = "pct_") %>% 
  filter(!is.na(change_group)) %>% 
  ggplot(., aes(change_group, pct)) +
  geom_boxplot() +
  facet_wrap(~race) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of racial demographics by change group",
       subtitle = str_wrap("These boxplots represent the median (bold line), 25th percentile (bottom of box), 75th percentile (top of box), and outliers (dots).", 100),
       x = "",
       y = "Number of schools")
```

```{r, eval = FALSE, include = FALSE}
# viz 2 - by high/low demographics
clean %>% 
  select(change_group, pct_white, pct_black, pct_hispanic, pct_aapi, pct_aian, pct_multiracial) %>% 
  mutate(pct_bipoc = 1-pct_white,
         pct_blackhispanic = pct_black+pct_hispanic) %>% 
  pivot_longer(cols = !change_group,
               names_to = "race",
               values_to = "pct",
               names_prefix = "pct_") %>% 
  filter(!is.na(change_group)) %>% 
  group_by(race) %>% 
  mutate(group = ifelse(pct > mean(pct, na.rm = TRUE), "Above average", "Below average")) %>% 
  ungroup() %>% 
  filter(!is.na(group)) %>% 
  group_by(change_group, race) %>% 
  count(group) %>% 
  ungroup() %>% 
  group_by(change_group, race) %>% 
  mutate(pct = n/sum(n),
         race = case_when(
    race == "aapi" ~ "Asian or Pacific Islander",
    race == "bipoc" ~ "BIPOC",
    race == "black" ~ "Black",
    race == "hispanic" ~ "Hispanic",
    race == "blackhispanic" ~ "Black or Hispanic",
    race == "multiracial" ~ "Multiracial",
    race == "white" ~ "White"
  ),
  pct = ifelse(group == "Below average", -1*pct, pct)) %>% 
  ggplot(., aes(pct, change_group, fill = group)) +
  geom_col() +
  facet_wrap(~race) +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  scale_fill_manual(values = c("#1F916F", "#e67f83")) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Distribution of districts with racial demographics above or below the sample average",
       subtitle = "Red dashed line indicates zero change.",
       x = "Percent of districts",
       y = "",
       fill = "Group") +
  geom_text(
  aes(
    label = scales::percent(abs(pct), accuracy = 1)
    #hjust = ifelse(group == "Below average", .05, -0.7)
  ),
  position = position_stack(vjust = .75),
  size = 3,
  show.legend = FALSE
) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

There is not strong statistical evidence that racial demographics differ by change group.

```{r}
mod <- clean %>% 
  mutate(pct_bipoc = 1-pct_white,
         pct_blackhispanic = pct_black+pct_hispanic)
mod1 <- lm(pct_aapi ~ change_group, data = mod)
mod2 <- lm(pct_aian ~ change_group, data = mod)
mod3 <- lm(pct_bipoc ~ change_group, data = mod)
mod4 <- lm(pct_black ~ change_group, data = mod)
mod5 <- lm(pct_hispanic ~ change_group, data = mod)
mod6 <- lm(pct_blackhispanic ~ change_group, data = mod)
mod7 <- lm(pct_white ~ change_group, data = mod)
modelsummary::modelsummary(
  models = list(
    "% AAPI" = mod1,
    "% AIAN" = mod2,
    "% BIPOC" = mod3,
    "% Black" = mod4,
    "% Hispanic" = mod5,
    "% Black & Hispanic" = mod6,
    "% White" = mod7
  )
)

# clean up
rm(mod, mod1, mod2, mod3, mod4, mod5, mod6, mod7)
```

## By % low income

We have data from 296 districts (3 missing) for the percentage of students considered low income by eligibility for free and reduced-price lunch. Overall descriptives are in the table below:

```{r}
# overall descriptives
psych::describe(clean$pct_lowinc) %>% 
  select(n, mean, sd, min, max) %>% 
  kable()
```

When we break this out by change group there is a slight pattern of increasing low-income students by number of domains adopted, but the difference is quite small.

```{r}
lowinc <- cont_table(clean, change_group, pct_lowinc, c(mean, sd, min, max), "A6.2 Distribution of the percentage of low-income students in a district by change group")
```

```{r}
# viz
cont_boxplot(clean, change_group, pct_lowinc,
             title = "Distribution of low-income students by change group",
             x_lab = "", y_lab = "Percent low-income")
```

```{r}
# alt viz
clean %>% 
  group_by(change_group) %>% 
  #count(num_schools) %>% 
  mutate(pct_lowinc_group = case_when(
    pct_lowinc < .25 ~ "Fewer than 25%",
    pct_lowinc >= .25 & pct_lowinc < .5 ~ "25-49%",
    pct_lowinc >= .5 & pct_lowinc < .75 ~ "50-75%",
    pct_lowinc >= .75 ~ "More than 75%"
  ),
  pct_lowinc_group = factor(pct_lowinc_group, levels = c("More than 75%", "50-75%", "25-49%", "Fewer than 25%"))) %>% 
  count(pct_lowinc_group) %>% 
  filter(!is.na(pct_lowinc_group)) %>% 
  ggplot(., aes(n, pct_lowinc_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 2),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       x = "Number of districts",
       y = "Number of schools in district")
```

There is not strong statistical evidence that the change group differs by the percentage of low-income students in the district (p ≈ .06).

```{r}
summary(lm(pct_lowinc ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
write.csv(lowinc, "output/tables/table_a6.2.csv", row.names = FALSE)
rm(lowinc)
```

## By % EL 

We have data from 296 districts (3 missing) for the percentage of students currently classified as English learners. Overall descriptives are in the table below:

```{r}
# overall descriptives
psych::describe(clean$pct_el) %>% 
  select(n, mean, sd, min, max) %>% 
  kable()
```

When we break this out by change group we see that the average percentage of EL-designated students in districts adopting more change (6-9 domains) is higher. Districts that are not currently working on any domains had a much smaller range--no districts had more than 21% EL-designated students.

```{r}
# generating descriptives by group
el <- cont_table(clean, change_group, pct_el, c(mean, sd, min, max), "A6.3 Distribution of the percentage of students designated as ELs in a district by change group")
```

```{r}
# viz
cont_boxplot(clean, change_group, pct_el, 
             title = "Distribution of EL-designated students by change group",
             x_lab = "", y_lab = "Percent EL")
```

```{r}
# alt viz
clean %>% 
  group_by(change_group) %>% 
  #count(num_schools) %>% 
  mutate(pct_el_group = case_when(
    pct_el < .1 ~ "Fewer than 10%",
    pct_el >= .1 & pct_el < .19 ~ "10-19%",
    pct_el >= .2 & pct_el < .3 ~ "20-29%",
    pct_el >= .3 ~ "More than 30%"
  ),
  pct_el_group = factor(pct_el_group, levels = c("More than 30%", "20-29%", "10-19%", "Fewer than 10%"))) %>% 
  count(pct_el_group) %>% 
  filter(!is.na(pct_el_group)) %>% 
  ggplot(., aes(n, pct_el_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 2),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       x = "Number of districts",
       y = "Number of schools in district")
```

There is not strong statistical evidence that the change group differs by the percentage of EL-designated students in the district (p ≈ .29).

```{r}
summary(lm(pct_el ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
write.csv(el, "output/tables/table_a6.3.csv", row.names = FALSE)
rm(el)
```

## By % SPED

We have data from 290 districts (9 missing) for the percentage of students with disabilities who are receiving special education services. Overall descriptives are in the table below:

```{r}
# overall descriptives
psych::describe(clean$pct_sped) %>% 
  select(n, mean, sd, min, max) %>% 
  kable()
```

When we break this out by change group we don't see notable differences in means or range.

```{r}
# generating descriptives by group
sped <- cont_table(clean, change_group, pct_sped, c(mean, sd, min, max), "A6.4 Distribution of the percentage of students with disabilities in a district by change group")
```

```{r}
# viz
cont_boxplot(clean, change_group, pct_sped,
             title = "Distribution of students with disabilities by change group",
             x_lab = "", y_lab = "Percent SWD")
```

```{r}
# alt viz
clean %>% 
  group_by(change_group) %>% 
  #count(num_schools) %>% 
  mutate(pct_sped_group = case_when(
    pct_sped < .1 ~ "Fewer than 10%",
    pct_sped >= .1 & pct_sped < .19 ~ "10-19%",
    pct_sped >= .2 & pct_sped < .3 ~ "20-29%",
    pct_sped >= .3 ~ "More than 30%"
  ),
  pct_sped_group = factor(pct_sped_group, levels = c("More than 30%", "20-29%", "10-19%", "Fewer than 10%"))) %>% 
  count(pct_sped_group) %>% 
  filter(!is.na(pct_sped_group)) %>% 
  ggplot(., aes(n, pct_sped_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 2),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       x = "Number of districts",
       y = "Number of schools in district")
```

There is evidence that the change group differs meaningfully by the percentage of students with disabilities in the district (p < .01). Notably, our tails are cut off for change groups with 0 domains or 6-9 domains. We may consider collapsing categories if we feel that this is an important category to look further into.

```{r}
summary(lm(pct_sped ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
write.csv(sped, "output/tables/table_a6.4.csv", row.names = FALSE)
rm(sped)
```

# A7 By student outcomes

## Math achievement 3-8

**UPDATE: This section has been re-run with centered variable 12.30.25**  

We have data on the percentage of students reaching proficiency in math across grades 3-8 for 266 districts (33 missing). The descriptives for the full sample are presented in the table below:

```{r}
# overall descriptives
psych::describe(clean$pct_math_prof_3_8_center)
```

When we break out descriptives by change group we see math achievement go down as the number of domains a district is working on increases.

```{r}
pct_math <- cont_table(clean, change_group, pct_math_prof_3_8_center, c(mean, sd, min, max), "A7.1 Math proficiency grades 3-8 by change group")
```

```{r}
cont_boxplot(clean, change_group, pct_math_prof_3_8_center, 
             title = "Math proficiency grades 3-8 by change group",
             x_lab = "",
             y_lab = "Percent proficient in math")
```

Change groups do seem to differ meaningfully by math proficiency with the original variable (p < .05). There is statistically lower proficiency rates for districts with more change (3-5 domains or 6-9 domains). However, this difference goes away when using the centered version (p = .09)

```{r}
summary(lm(pct_math_prof_3_8_center ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(pct_math, "output/tables/table_a7.1.csv", row.names = FALSE)
rm(pct_math)
```

## Math achievement 9-12

We have data on the percentage of students reaching proficiency in math across grades 9-12 for 187 districts (112 missing). The descriptives for the full sample are presented in the table below:

```{r}
# overall descriptives
psych::describe(clean$pct_math_prof_9_12_center)
```

When we break out by change group we see the same pattern holds at the secondary level; average math achievement is lower as the number of domains a district is working on increases.

```{r}
pct_math <- pct_math <- cont_table(clean, change_group, pct_math_prof_9_12_center, c(mean, sd, min, max), "A7.2 Math proficiency grades 9-12 by change group")
```

```{r}
cont_boxplot(clean, change_group, pct_math_prof_9_12_center, 
             title = "Math proficiency grades 9-12 by change group",
             x_lab = "",
             y_lab = "Percent proficient in math")
```

Change groups do not differ meaningfully by math proficiency (p = .2).

```{r}
summary(lm(pct_math_prof_9_12_center ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(pct_math, "output/tables/table_a7.2.csv", row.names = FALSE)
rm(pct_math)
```

## Reading achievement 3-8

We have data on the percentage of students reaching proficiency in ELA across grades 3-8 for 266 districts (33 missing). The descriptives for the full sample are presented in the table below:

```{r}
# overall descriptives
psych::describe(clean$pct_ela_prof_3_8_center)
```

When we break out by change group we see the same pattern holds for ELA at the primary level; average reading achievement is lower as the number of domains a district is working on increases, though the magnitude is less severe than it is for math.

```{r}
pct_ela <- cont_table(clean, change_group, pct_ela_prof_3_8_center, c(mean, sd, min, max), "A7.3 ELA proficiency grades 3-8 by change group")
```

```{r}
cont_boxplot(clean, change_group, pct_ela_prof_3_8_center, 
             title = "ELA proficiency grades 3-8 by change group",
             x_lab = "",
             y_lab = "Percent proficient in reading")
```

Like with math proficiency in earlier grades, change groups did seem to differ meaningfully by reading proficiency (p < .05). Specifically, districts with more change (3-5 domains and 6-9 domains) are statistically different from those with less change. However, when we use the centered version that difference goes away (p = .12).

```{r}
summary(lm(pct_ela_prof_3_8_center ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(pct_ela, "output/tables/table_a7.3.csv", row.names = FALSE)
rm(pct_ela)
```

## Reading achievement 9-12

We have data on the percentage of students reaching proficiency in ELA across grades 9-12 for 187 districts (112 missing). The descriptives for the full sample are presented in the table below:

```{r}
# overall descriptives
psych::describe(clean$pct_ela_prof_9_12_center)
```

The same pattern again holds for reading achievement at the secondary level.

```{r}
pct_ela <- cont_table(clean, change_group, pct_ela_prof_9_12_center, c(mean, sd, min, max), "A7.4 ELA proficiency grades 9-12 by change group")
```

```{r}
cont_boxplot(clean, change_group, pct_ela_prof_9_12_center, 
             title = "ELA proficiency grades 9-12 by change group",
             x_lab = "",
             y_lab = "Percent proficient in reading")
```

Change groups do seem to differ meaningfully by reading proficiency for secondary students, regardless of the specification (p < .05).

```{r}
summary(lm(pct_ela_prof_9_12_center ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(pct_ela, "output/tables/table_a7.4.csv", row.names = FALSE)
rm(pct_ela)
```

# A8 Chronic absenteeism

We have chronic absenteeism data for 253 districts (46 missing). We do not have any data from New Hampshire for this variable (they have a retention variable that is not comparable which we could look at instead to test patterns).

```{r}
#overview
psych::describe(clean$pct_chronic_absenteeism)
```

There are no obvious patterns when we break out chronic absenteeism by change group, though it may be worth noting that the range for districts focusing on 1 or 2 domains is much larger (up to 95%). Rates go up by a hair as a district works on more domains - but the difference is so small this may not be statistically significant.

```{r}
pct_absent <- cont_table(clean, change_group, pct_chronic_absenteeism, c(mean, sd, min, max), "A.8 Chronic absenteeism rates by change group")
```

```{r}
cont_boxplot(clean, change_group, pct_chronic_absenteeism,
             title = "Chronic absenteeism rates by change group",
             x_lab = "",
             y_lab = "Percent chronically absent")
```

Change groups do not differ meaningfully by absenteeism rates (p = .39).

```{r}
summary(lm(pct_chronic_absenteeism ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(pct_absent, "output/tables/table_a8.csv", row.names = FALSE)
rm(pct_absent)
```


# A9 4-year graduation

We have 4-year graduation rates only among the 230 districts who offer secondary grades (69 missing). The descriptives across the full sample are in the table below:

```{r}
# overview
psych::describe(clean$grad_rate)
```

Graduation rates are slightly lower the more domains a district is working on.

```{r}
grad_rate <- cont_table(clean, change_group, grad_rate, c(mean, sd, min, max), "A.9 4-year graduation rates by change group")
```

```{r}
cont_boxplot(clean, change_group, grad_rate,
             title = "4-year graduation rates by change group",
             x_lab = "",
             y_lab = "Percent graduating")
```

The difference in means across change groups is not statistically significant by absenteeism rates, though the effect is marginal (p = .06).

```{r}
summary(lm(grad_rate ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
# save
write.csv(grad_rate, "output/tables/table_a9.csv", row.names = FALSE)
rm(grad_rate)
```


# A10 Other

## Retention rate

We have retention data for 182 districts in Connecticut, Massachusetts, and Vermont (117 missing; missing entirely for Maine, New Hampshire, and Rhode Island). The descriptive statistics across the subsestted sample are presented in the table below: 

```{r}
#overview
psych::describe(clean$retention_rate)
```

There are no obvious patterns when we break out retention rates by change group.

```{r}
retention <- cont_table(clean, change_group, retention_rate, c(mean, sd, min, max), "A10.1 Retention rates by change group")
```

```{r}
cont_boxplot(clean, change_group, retention_rate,
             title = "Retention rates by change group",
             x_lab = "",
             y_lab = "Percent retained")
```

Change groups do not differ meaningfully by retention rates (p = .82).

```{r}
summary(lm(retention_rate ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(retention, "output/tables/table_a10.1.csv", row.names = FALSE)
rm(retention)
```

## Per-pupil expenditure

We have per-pupil expenditure data for 294 districts (5 missing). The descriptive statistics across the sample are presented in the table below: 

```{r}
#overview
psych::describe(clean$per_pupil_spending)
```

Spending seems to be roughly similar across change groups, though the standard deviation gets much smaller as the number of domains a district is working on increases. For districts with 6-9 domains, the standard deviation is only $3,349 compared to $6,077 for districts with 0 domains. This implies there's less variation in spending for the former, while districts may vary much more among other groups with less change.

```{r}
spending <- cont_table(clean, change_group, per_pupil_spending, c(mean, sd, min, max), "A10.2 Per-pupil expenditure by change group")
```

```{r}
cont_boxplot(clean, change_group, per_pupil_spending,
             title = "Per pupil spending by change group",
             x_lab = "",
             y_lab = "Per pupil spending ($)")
```

Change groups do not differ meaningfully by spending (p = .64).

```{r}
summary(lm(per_pupil_spending ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(spending, "output/tables/table_a10.2.csv", row.names = FALSE)
rm(spending)
```

## Teacher certification

We have teacher certification data for 285 districts (14 missing). The descriptive statistics across the sample are presented in the table below: 

```{r}
#overview
psych::describe(clean$pct_fte_cert)
```

Teacher certification seems to be roughly similar across change groups.

```{r}
fte <- cont_table(clean, change_group, pct_fte_cert, c(mean, sd, min, max), "A10.3 Percent of certified teachers (FTE) by change group")
```

```{r}
cont_boxplot(clean, change_group, pct_fte_cert,
             title = "Percent of certified teachers (FTE) by change group",
             x_lab = "",
             y_lab = "Percent certified (FTE)")
```

Change groups do not differ meaningfully by teacher certification (p = .62).

```{r}
summary(lm(pct_fte_cert ~ change_group, data = clean))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(fte, "output/tables/table_a10.3.csv", row.names = FALSE)
rm(fte)
```

## Percent regularly attending

Since New Hampshire was missing both retention data and chronic absenteeism data, we are using an alternate measure to test the larger trend: percentage of students regularly attending. This variable was not able to be transformed to chronic absenteeism. The descriptive statistics for New Hampshire districts are presented in the table below:

```{r}
nh <- clean %>% filter(state == "NH")
psych::describe(nh$pct_reg_attend)
```

Teacher certification seems to be roughly similar across change groups.

```{r}
reg_attend <- cont_table(nh, change_group, pct_reg_attend, c(mean, sd, min, max), "A10.4 Percent of students regularly attending by change group in New Hampshire")
```

```{r}
cont_boxplot(nh, change_group, pct_reg_attend,
             title = "Percent of students regularly attending by change group in New Hampshire",
             x_lab = "",
             y_lab = "Percent regularly attending")
```

Change groups do not differ meaningfully by teacher certification (p = .7).

```{r}
summary(lm(pct_reg_attend ~ change_group, data = nh))
```

```{r, include = FALSE, eval = FALSE}
#save
write.csv(reg_attend, "output/tables/table_a10.4.csv", row.names = FALSE)
rm(reg_attend)
```