---
title: "Phase II - District Characteristics by Change Group"
author: "Janette Avelar"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: flatly
    highlight: tango
  word_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # hides code
  message = FALSE, # hides messages
  warning = FALSE,  # hides warnings
  fig.width = 10, fig.height = 9
)
library(rio)
library(tidyverse)
library(DT)
library(kableExtra)

dat <- import('/Users/janetteavelar/Library/CloudStorage/GoogleDrive-javelar3@asu.edu/.shortcut-targets-by-id/1NdfrCmhAW0NpUV47ruVeKThD8M3arhRq/DATA: Barr 3.0/Survey/Analysis/Current Data Files/MergedSurvey_251215.dta')
```

# Survey Task

In this analysis, we are interested in the guiding question: **How does context relate to change groups?**  

There are 4 change groups, defined as:  
- 0 changes underway  
- 1-2 changes underway  
- 3-5 changes underway  
- 6-9 changes underway  

This file looks at district characteristics, comparing the mean and distribution of state, district size (enrollment), district size (number of schools), serve high school, enrollment change, racial composition (% White), % low inc, % EL % SWD, student outcomes by change group.  

**Note**: *There are 8 districts that do not fall into a change group and I'm not totally sure why -- I checked their responses for Q4-6, and 2 of the districts had responded to all 3, while some seem due to missingness. Would be good to talk about who these districts are and why they're missing.*  

```{r}
# overview
dat %>% 
  filter(is.na(change_group)) %>% 
  select(qcode, partial, state, total_domains_working,change_group) %>% 
  kable()

# q6
tmp <- dat %>%
  filter(is.na(change_group)) %>%
  select(qcode, starts_with("q6"))

highlight <- which(!is.na(tmp$q6_2))

tmp %>%
  kable() %>%
  kable_styling() %>%
  row_spec(highlight, color = "black", background = "yellow")
rm(tmp, highlight)
```


```{r}
# Clean up data
clean <- dat %>% 
  haven::zap_label() %>% 
  mutate(grades_served = case_when(
    grades_served == 1 ~ "K-6",
    grades_served == 2 ~ "K-8",
    grades_served == 3 ~ "K-12",
    grades_served == 4 ~ "6-8",
    grades_served == 5 ~ "7-12",
    grades_served == 6 ~ "9-12"
  ),
  district_type = case_when(
    district_type == 1 ~ "Traditional",
    district_type == 2 ~ "Regional",
    district_type == 3 ~ "Regional CTE",
    district_type == 4 ~ "Charter",
    district_type == 5 ~ "State agency"
  ),
  urbanicity = case_when(
    urbanicity == 1 ~ "Urban",
    urbanicity == 2 ~ "Suburban",
    urbanicity == 3 ~ "Rural"
  ),
  ma_account_code = case_when(
    ma_account_code == 1 ~ "Insufficient data",
    ma_account_code == 2 ~ "Not requiring assistance",
    ma_account_code == 3 ~ "Requiring assistance or intervention"
  ),
  change_group = case_when(
    change_group == 1 ~ "0 domains",
    change_group == 2 ~ "1-2 domains",
    change_group == 3 ~ "3-5 domains",
    change_group == 4 ~ "6-9 domains"
  ),
  serve_hs = case_when(
    serve_hs == 0 ~ "Does not serve high school",
    serve_hs == 1 ~ "Serves high school"
  ))
```

# A1 By state

There are 6 states in our data:  

```{r}
table(clean$state) %>% kable()
```

The percentages in this table represent the distribution within a given state, e.g., *"In Connecticut, 18% of districts were not undergoing change (0 domains) while 6% had high change (6-9 domains)."*  

The pattern of distributions is fairly even, but there are some differences to note:  
- Maine and Rhode Island both are skewed toward less, or more focused, change, with the majority (42% & 47% respectively) of districts in the state focusing on 1 or 2 changes.  
- New Hampshire had the highest proportion of districts not focusing on change at all (0 domains - 20%) and the smallest proportion of districts undergoing a lot of change (6-9 domains - 2%).  
- Vermont had the smallest proportion of districts undergoing no change (0 domains - 8%).  

```{r}
# generate counts
state <- clean %>% 
  group_by(change_group) %>% 
  count(state) %>% 
  ungroup() %>% 
  group_by(state) %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  # Note: rounding to 2 decimals was equivalent when testing pct overall or pct within group
  mutate(change_group = ifelse(is.na(change_group), "Missing", change_group),
         pct = round(n/total, 2),
         state = case_when(
           state == "CT" ~ "Connecticut",
           state == "MA" ~ "Massachusetts",
           state == "ME" ~ "Maine",
           state == "NH" ~ "New Hampshire",
           state == "RI" ~ "Rhode Island",
           state == "VT" ~ "Vermont"
         )) %>% 
  select(change_group, state, pct) %>% 
  pivot_wider(names_from = change_group,
              values_from = pct)

# table
datatable(state,
          caption = "Table A1 Change groups by state",
          options = list(searchHighlight = TRUE))

# viz
state %>% 
  pivot_longer(cols = !state,
               names_to = "change_group",
               values_to = "pct") %>% 
  filter(change_group != "Missing") %>% 
  mutate(change_group = factor(change_group, levels = c("6-9 domains", "3-5 domains", "1-2 domains", "0 domains"))) %>% 
  ggplot(., aes(pct, change_group)) +
  geom_col(fill = "#1F916F") +
  scale_x_continuous(expand = c(0, 0),
                     limits = c(0, 1)) +
  scale_y_discrete(drop = FALSE) +
  theme_bw() +
  labs(title = "Change groups by state",
       x = "Percent",
       y = "") +
  geom_text(
      aes(label = scales::percent(pct, accuracy = 1)),
      position = position_nudge(x = 0.08),
      size = 3,
      show.legend = FALSE
    ) +
  facet_wrap(~state) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```

```{r include = FALSE, eval = FALSE}
# save table
write.csv(state, "output/tables/table_a1.csv", row.names = FALSE)
rm(state)
```

# A2 By district size

## Defined as enrollment

We have enrollment information for the 2024-25 school across 304 districts. The table below includes descriptive statistics across the broader sample:

```{r}
# overall distribution
psych::describe(clean$enrollment_2025) %>% 
  select(n, mean, sd, min, max) %>% 
  kable()
```

Table A2.1 displays descriptive statistics broken out by the number of domains a given district is working on.

```{r}
enroll <- clean %>% 
  group_by(change_group) %>% 
  summarize(mean_enroll = mean(enrollment_2025, na.rm = TRUE),
            sd_enroll = sd(enrollment_2025, na.rm = TRUE),
            min_enroll = min(enrollment_2025, na.rm = TRUE),
            max_enroll = max(enrollment_2025, na.rm = TRUE)) %>% 
  mutate(change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  mutate(across(c(mean_enroll, sd_enroll), ~round(., 0)))

# table
datatable(enroll, 
          caption = "A2.1 Enrollment by Change Group")

# viz
ggplot(clean, aes(change_group, enrollment_2025)) +
  geom_boxplot() +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(title = "Distribution of enrollment in 2025 by change group",
       subtitle = str_wrap("These boxplots represent the median (bold line), 25th percentile (bottom of box), 75th percentile (top of box), and outliers (dots).", 100),
       x = "",
       y = "Enrollment")

# alt viz
# SHOULD I ADD PERCENTAGES???
# If yes, need to think about denominator - do we want it to be among domain group, or among enrollment band?
clean %>% 
  mutate(enroll_group = case_when(
    enrollment_2025 < 500 ~ "< 500",
    enrollment_2025 >= 500 & enrollment_2025 < 1500 ~ "500-1,499",
    enrollment_2025 >= 1500 & enrollment_2025 < 3000 ~ "1,500-2,999",
    enrollment_2025 >= 3000 & enrollment_2025 < 7500 ~ "3,000-7,499",
    #enrollment_2025 >= 5000 & enrollment_2025 < 7500 ~ "5,000-7,500",
    enrollment_2025 >= 7500 ~ "> 7,500"
  )) %>% 
  group_by(change_group) %>% 
  count(enroll_group) %>% 
  filter(!is.na(enroll_group)) %>% 
  filter(!is.na(change_group)) %>% 
  mutate(enroll_group = factor(enroll_group, levels = c("> 7,500", "3,000-7,499", "1,500-2,999", "500-1,499", "< 500"))) %>% 
  ggplot(., aes(n, enroll_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0),
                     limits = c(0, 50)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 1.5),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of enrollment size by change group",
       x = "Number of districts",
       y = "Number of students")
```

```{r, include = FALSE, eval = FALSE}
# save table
write.csv(enroll, "output/tables/table_a2.1.csv", row.names = FALSE)
rm(enroll)
```


## Defined as number of schools

The table below includes descriptive statistics for the number of schools and number of high schools districts have:

```{r}
# overall descriptives
subset <- clean %>% 
  select(num_schools, num_high_schools)
psych::describe(subset) %>% 
  select(n, mean, sd, min, max) %>% 
  kable()
rm(subset)
```

```{r}
# generating descriptives by group
n_school <- clean %>% 
  group_by(change_group) %>% 
  summarize(mean_high_schools = mean(num_high_schools, na.rm = TRUE),
            sd_high_schools = sd(num_high_schools, na.rm = TRUE),
            min_high_schools = min(num_high_schools, na.rm = TRUE),
            max_high_schools = max(num_high_schools, na.rm = TRUE),
            mean_all_schools = mean(num_schools, na.rm = TRUE),
            sd_all_schools = sd(num_schools, na.rm = TRUE),
            min_all_schools = min(num_schools, na.rm = TRUE),
            max_all_schools = max(num_schools, na.rm = TRUE)) %>% 
  pivot_longer(cols = -change_group,
               names_to = c("stat", "group"),
               names_pattern = "(mean|sd|min|max)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat,
              values_from = value) %>%
  mutate(group = recode(group,
                        high_schools = "High school",
                        all_schools  = "All schools"),
         change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  mutate(across(c(mean, sd), ~round(., 2)))

# table
datatable(n_school,
          caption = "A2.2 Distribution of number of total schools and high schools by change group")

# viz
clean %>% 
  select(change_group, num_schools, num_high_schools) %>% 
  pivot_longer(cols = c("num_schools", "num_high_schools"),
               names_to = "group",
               values_to = "n_school",
               names_prefix = "num_") %>% 
  mutate(group = ifelse(group == "schools", "All schools", "High schools"),
         change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  ggplot(., aes(change_group, n_school)) +
  geom_boxplot() +
  facet_wrap(~group) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       subtitle = str_wrap("These boxplots represent the median (bold line), 25th percentile (bottom of box), 75th percentile (top of box), and outliers (dots).", 100),
       x = "",
       y = "Number of schools")

# alt viz
clean %>% 
  group_by(change_group) %>% 
  #count(num_schools) %>% 
  mutate(num_schools_group = case_when(
    num_schools < 5 ~ "< 5",
    num_schools >= 5 & num_schools < 10 ~ "5-9",
    num_schools >= 10 & num_schools < 15 ~ "10-14",
    num_schools >= 15 & num_schools < 20 ~ "15-19",
    num_schools >= 20 & num_schools < 25 ~ "20-24",
    num_schools >= 25 ~ "25+"
  ),
  num_schools_group = factor(num_schools_group, levels = c("25+", "20-24", "15-19", "10-14", "5-9", "< 5"))) %>% 
  count(num_schools_group) %>% 
  filter(!is.na(change_group)) %>% 
  ggplot(., aes(n, num_schools_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 2),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       x = "Number of districts",
       y = "Number of schools in district")
```

```{r, include = FALSE, eval = FALSE}
write.csv(n_school, "output/tables/table_a2.2.csv", row.names = FALSE)
```


# A3 By serving high school

```{r}
# alt viz
clean %>% 
  group_by(change_group) %>% 
  #count(num_high_schools) 
  mutate(num_schools_group = case_when(
    num_high_schools == 0 ~ "0",
    num_high_schools == 1 ~ "1",
    num_high_schools == 2 ~ "2",
    num_high_schools == 3 ~ "3",
    num_high_schools >= 4 ~ "4+"
  ),
  num_schools_group = factor(num_schools_group, levels = c("4+", "3", "2", "1", "0"))) %>% 
  count(num_schools_group) %>% 
  filter(!is.na(change_group)) %>% 
  ggplot(., aes(n, num_schools_group)) +
  geom_col(fill = "#1F916F") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~change_group) +
  geom_text(
      aes(label = n),
      position = position_nudge(x = 2),
      size = 3.5,
      show.legend = FALSE) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of number of schools by change group",
       x = "Number of districts",
       y = "Number of high schools in district")
```

```{r}
# clean up
rm(n_school)
```


# A4 By enrollment change

We have 6-year enrollment data for 296 districts (11 missing) indicating enrollment change over that time. Enrollment is trending downwards in our data, with the average district dropping 3 percentage points (check interpretation) between 2019 and 2025. The majority of districts (74%) had a value below 0, indicating a loss in enrollment.

```{r}
# overall descriptives
psych::describe(clean$enrollment_change_6yr_pct)
# 26% gaining, 74% losing; 11 districts held missing values
clean %>% 
  mutate(group = ifelse(enrollment_change_6yr_pct <= 0, "Loss", "Gain")) %>% 
  #count(group)
  ggplot(., aes(enrollment_change_6yr_pct)) +
  geom_histogram(fill = "#1F916F") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  theme_bw() +
  labs(title = "Distribution of 6-year enrollment change across our sample",
       subtitle = "Red dashed line indicates zero change.",
       x = "6-year enrollment change",
       y = "Number of districts")
```

When we break out enrollment change by change group, we don't see any obvious differences in the pattern of loss. Districts with a higher degree of change (6-9 domains) are experiencing a slightly higher rate of loss compared to other groups (6%).

```{r}
# generating descriptives by group
enroll_change <- clean %>% 
  group_by(change_group) %>% 
  summarize(mean = mean(enrollment_change_6yr_pct, na.rm = TRUE),
            sd = sd(enrollment_change_6yr_pct, na.rm = TRUE),
            min = min(enrollment_change_6yr_pct, na.rm = TRUE),
            max = max(enrollment_change_6yr_pct, na.rm = TRUE)) %>% 
  mutate(change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  mutate(across(c(mean, sd, min, max), ~round(., 2)))

# table
datatable(enroll_change,
          caption = "A4 Distribution of enrollment change by change group")

# first viz
clean %>% 
  filter(!is.na(change_group)) %>% 
  mutate(group = ifelse(enrollment_change_6yr_pct <= 0, "Loss", "Gain")) %>% 
  #count(group)
  ggplot(., aes(enrollment_change_6yr_pct)) +
  geom_histogram(fill = "#1F916F") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  theme_bw() +
  labs(title = "Distribution of 6-year enrollment change among change groups",
       subtitle = "Red dashed line indicates zero change.",
       x = "6-year enrollment change",
       y = "Number of districts") +
  facet_wrap(~change_group)

# second viz
clean %>% 
  filter(!is.na(change_group)) %>% 
  mutate(group = ifelse(enrollment_change_6yr_pct <= 0, "Loss", "Gain")) %>% 
  group_by(change_group) %>% 
  filter(!is.na(group)) %>% 
  count(group) %>% 
  mutate(pct = n/sum(n),
         pct = ifelse(group == "Loss", pct*-1, pct)) %>% 
  ggplot(., aes(pct, change_group, fill = group)) +
  geom_col() +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  scale_fill_manual(values = c("#1F916F", "#e67f83")) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Distribution of districts with enrollment gains or loss between 2019 and 2025 across change groups",
       subtitle = "Red dashed line indicates zero change.",
       x = "Percent of districts",
       y = "",
       fill = "Enrollment trend") +
  geom_text(
  aes(label = scales::percent(abs(pct), accuracy = 1),
      hjust = ifelse(group == "Loss", 1.05, -0.05)),
  position = position_stack(vjust = .5),
  size = 3.5,
  show.legend = FALSE) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

```{r, include = FALSE, eval = FALSE}
write.csv(enroll_change, "output/tables/table_a4.csv", row.names = FALSE)
rm(enroll_change)
```


# A5 By racial composition

We have racial demographics data for about 304 districts (3 missing overall, 4 missing American Indian and Alaska Native data), including:  
- % White students  
- % Black students  
- % Latine/Hispanic students  
- % Asian or Pacific Islander students  
- % American Indian or Alaska Native  
- % Multiracial  

The overall descriptives for our sample are in the table below:

```{r}
# overall descriptives
clean %>% 
  select(pct_white, pct_black, pct_hispanic, pct_aapi, pct_aian, pct_multiracial) %>% 
  psych::describe()
```

When we break out descriptives by change group we see that districts undergoing more change (6-9 domains) tend to have higher populations of BIPOC students and combined Black/Hispanic student populations.  

```{r}
race <- clean %>% 
  mutate(bipoc = 1-pct_white,
         blackhispanic = pct_black+pct_hispanic) %>% 
  group_by(change_group) %>% 
  summarize(mean_white = mean(pct_white, na.rm = TRUE),
            sd_white = sd(pct_white, na.rm = TRUE),
            mean_black = mean(pct_black, na.rm = TRUE),
            sd_black = sd(pct_black, na.rm = TRUE),
            mean_hispanic = mean(pct_hispanic, na.rm = TRUE),
            sd_hispanic = sd(pct_hispanic, na.rm = TRUE),
            mean_aapi = mean(pct_aapi, na.rm = TRUE),
            sd_aapi = sd(pct_aapi, na.rm = TRUE),
            mean_aian = mean(pct_aian, na.rm = TRUE),
            sd_aian = sd(pct_aian, na.rm = TRUE),
            mean_multiracial = mean(pct_multiracial, na.rm = TRUE),
            sd_multiracial = sd(pct_multiracial, na.rm = TRUE),
            mean_bipoc = mean(bipoc, na.rm = TRUE),
            sd_bipoc = sd(bipoc, na.rm = TRUE),
            mean_blackhispanic = mean(blackhispanic, na.rm = TRUE),
            sd_blackhispanic = sd(blackhispanic, na.rm = TRUE)) %>% 
  mutate(change_group = ifelse(is.na(change_group), "Missing", change_group)) %>% 
  pivot_longer(cols = !change_group,
               names_to = c("stat", "race"),
               names_pattern = "(mean|sd)_(.*)",
               values_to = "value") %>% 
  pivot_wider(names_from = stat,
              values_from = value) %>% 
  pivot_wider(names_from = change_group,
              values_from = c(mean, sd)) %>% 
  mutate(race = case_when(
    race == "white" ~ "White",
    race == "black" ~ "Black",
    race == "hispanic" ~ "Hispanic",
    race == "aapi" ~ "Asian or Pacific Islander",
    race == "aian" ~ "American Indian or Alaska Native",
    race == "multiracial" ~ "Multiracial",
    race == "bipoc" ~ "All BIPOC students",
    race == "blackhispanic" ~ "Combined Black + Hispanic"
  )) %>% 
  mutate(across(!race, ~round(., 2))) %>% 
  select(race, `mean_0 domains`, `sd_0 domains`, `mean_1-2 domains`, `sd_1-2 domains`, `mean_3-5 domains`, `sd_3-5 domains`, `mean_6-9 domains`, `sd_6-9 domains`) 
# display
sketch = htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(colspan = 1, ""),
      th(colspan = 2, '0 domains'),
      th(colspan = 2, '1-2 domains'),
      th(colspan = 2, '3-5 domains'),
      th(colspan = 2, "6-9 domains")
    ),
    tr(
      lapply(c("Group", rep(c('Mean', 'SD'), 4)), th)
    )
  )
))
datatable(race,
          caption = "A5 District racial demographics by change group",
          colnames = c("", rep(c("Mean", "SD"), 4)),
          container = sketch,
          rownames = FALSE)
```

```{r, include = FALSE, eval = FALSE}
write.csv(race, "output/tables/table_a5.csv", row.names = FALSE)
rm(race, sketch)
```

```{r}
# viz 1 - boxplots
clean %>% 
  select(change_group, pct_white, pct_black, pct_hispanic, pct_aapi, pct_aian, pct_multiracial) %>% 
  mutate(pct_bipoc = 1-pct_white,
         pct_blackhispanic = pct_black+pct_hispanic) %>% 
  pivot_longer(cols = !change_group,
               names_to = "race",
               values_to = "pct",
               names_prefix = "pct_") %>% 
  filter(!is.na(change_group)) %>% 
  ggplot(., aes(change_group, pct)) +
  geom_boxplot() +
  facet_wrap(~race) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of racial demographics by change group",
       subtitle = str_wrap("These boxplots represent the median (bold line), 25th percentile (bottom of box), 75th percentile (top of box), and outliers (dots).", 100),
       x = "",
       y = "Number of schools")
```

```{r, eval = FALSE, include = FALSE}
# viz 2 - by high/low demographics
clean %>% 
  select(change_group, pct_white, pct_black, pct_hispanic, pct_aapi, pct_aian, pct_multiracial) %>% 
  mutate(pct_bipoc = 1-pct_white,
         pct_blackhispanic = pct_black+pct_hispanic) %>% 
  pivot_longer(cols = !change_group,
               names_to = "race",
               values_to = "pct",
               names_prefix = "pct_") %>% 
  filter(!is.na(change_group)) %>% 
  group_by(race) %>% 
  mutate(group = ifelse(pct > mean(pct, na.rm = TRUE), "Above average", "Below average")) %>% 
  ungroup() %>% 
  filter(!is.na(group)) %>% 
  group_by(change_group, race) %>% 
  count(group) %>% 
  ungroup() %>% 
  group_by(change_group) %>% 
  mutate(pct = n/sum(n))
race = case_when(
    
  ))
  ggplot(., aes(change_group, pct, fill = group)) +
  geom_col() +
  facet_wrap(~race)
```


## % White (BIPOC?)

```{r}

```

## % Black & Latine

```{r}

```

## Specific groups?

```{r}

```

# By % low income

```{r}

```

# By % EL

```{r}

```

# By % SPED

```{r}

```

# By student outcomes

## Math achievement 3-8

```{r}

```

## Math achievement 9-12

```{r}

```

## Reading achievement 3-8

```{r}

```

## Reading achievement 9-12

```{r}

```

## Attendance

[NH I think only had chronic absenteeism? Will have to double check]

```{r}

```

## 4-year graduation

```{r}

```

## Retention rate