---
title: "Non-Prioritized Goals in NE Districts"
author: "Janette Avelar"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: flatly
    highlight: tango
    code_folding: hide
  word_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # hides code
  message = FALSE, # hides messages
  warning = FALSE,  # hides warnings
  fig.width = 10, fig.height = 9
)
library(here)
library(rio)
library(tidyverse)
library(DT)
library(kableExtra)
dat <- import('/Users/janetteavelar/Library/CloudStorage/GoogleDrive-javelar3@asu.edu/.shortcut-targets-by-id/1NdfrCmhAW0NpUV47ruVeKThD8M3arhRq/DATA: Barr 3.0/Survey/Analysis/JA/Barr_Survey_20251117.dta')
labs <- import(here("data/clean", "survey_labs.csv")) %>% 
  mutate(var = str_to_lower(var))
```

The goal of this analysis is to understand the goals district superintendents did *not* select as needing substantial change, and whether there are similarities across these districts that might help explain their selection patterns. The analysis focuses on questions 4 & 5:

**In order to achieve your district’s long-term goals, how much change is needed in the following areas related to teaching, learning, and student support?**

I'm specifically interested in the response option *no change required*, thus responses are coded to 1 and 0 for any other response (except `NA`).

```{r}
# Create vector for renaming columns
name_map <- setNames(labs$var, labs$var_name)

# Data prep
clean <- dat %>% 
  # strip labels
  haven::zap_label() %>% 
  # recode all missing values to NA
  mutate(across(where(is.numeric), ~na_if(., -99))) %>% 
  # select and rename relevant cols
  rename(any_of(name_map)) %>% 
  # swap cat
  mutate(supt_tenure = case_when(
    supt_tenure == 1 ~ "Less than 1 year",
    supt_tenure == 2 ~ "1–3 years",
    supt_tenure == 3 ~ "4–6 years",
    supt_tenure == 4 ~ "7–10 years",
    supt_tenure == 5 ~ "More than 10 years"
  )) %>% 
  # backwards code "need" questions (Q4 & 5) so 1 = not selected
  mutate(across(starts_with("need"), ~ifelse(. == 1, 1, 0))) %>% 
  # recode helps/hinders columns to cat
  mutate(across(starts_with("helps"), ~case_when(
    .x == 1 ~ "Definitely helps", 
    .x == 2 ~ "Helps more than hinders", 
    .x == 3 ~ "Hinders more then helps", 
    .x == 4 ~ "Definitely hinders" 
  ))) %>% 
  # recode personal opinion/belief questions to cat
  mutate(across(starts_with("op"), ~case_when(
    .x == 1 ~ "Definitely true",
    .x == 2 ~ "Mostly true",
    .x == 3 ~ "Mostly false",
    .x == 4 ~ "Definitely false"
  ))) 
# %>% 
#   # TEMPORARY ZOOMING IN
#   select(
#     -starts_with("need"),
#     any_of(c(
#       "need_support_adult_respect",
#       "need_system_budget_alignment",
#       "need_support_placement_structures"
#     ))
#   )
```

# Basic descriptives

```{r}
# create desc df
desc <- clean %>% 
  select(starts_with("need") & !ends_with("other")) %>% 
  psych::describe() %>% 
  rownames_to_column(var = "var_name") %>%
  select(-vars) %>% 
  left_join(labs, by = "var_name") %>% 
  select(var = nice_label, n, mean, sd, min, max) %>% 
  mutate(across(mean:max, ~round(., 2)))
# display table
datatable(desc,
          colnames = c("Goals not being prioritized", "Total respondents", "% Not prioritizing", "SD", "Min", "Max"))
```

```{r}
# Create a df of N sizes
n_tab <- clean %>% 
  select(starts_with("need") & !ends_with("other")) %>% 
  summarize(across(everything(), ~sum(., na.rm = TRUE))) %>% 
  pivot_longer(cols = everything(),
               names_to = "var_name",
               values_to = "n")
```

```{r}
# create viz
clean %>% 
  select(starts_with("need") & !ends_with("other")) %>% 
  pivot_longer(cols = everything(),
               names_to = "var_name",
               values_to = "selected") %>% 
  filter(!is.na(selected)) %>% 
  group_by(var_name, selected) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(var_name) %>% 
  mutate(total = sum(n),
         pct = n/total) %>% 
  filter(selected == 1) %>% 
  ungroup() %>% 
  left_join(labs, by = "var_name") %>% 
  mutate(nice_label = paste0(nice_label, " (N = ", n, ")")) %>% 
  select(nice_label, n, total, pct) %>% 
  ggplot(., aes(pct, reorder(nice_label, pct))) +
  geom_col(fill = "#004B72") +
  geom_text(
      aes(label = ifelse(pct > 0, scales::percent(pct, accuracy = 1), "")),
      hjust = 1.1,
      size = 4,
      color = "white"
    ) +
    scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 40)) +
    scale_x_continuous(expand = expansion(mult = c(0, .02)),
                       limits = c(0, .5)) +
    labs(x = "Percent indicating item does not need substantial change", 
         y = "", 
         title = stringr::str_wrap("Distribution of goals superintendents don't feel need substantial change", 65)) +
    theme_bw()
```


# Stacked bars

```{r}
# Pull in functions
source(here("functions/plot_non_priority.R"))
```

## By tenure

```{r}
not_priority_cat(
  clean,
  supt_tenure,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by tenure",
  recode_map = c("More than 10 years" = "10+ years"),
  var_levels = c("10+ years", "7–10 years", "4–6 years", "1–3 years", "Less than 1 year")
)
```

## By goal: Student engagement

```{r}
not_priority_binary(clean, goal_engagement, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected increasing student engagement as a long-term goal", n_tab)
```

## By goal: Create a safe and welcoming school community

```{r}
not_priority_binary(clean, goal_safe_community, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected creating a safe and welcoming school community as a long-term goal", n_tab)
```

## By goal: Improve academic achievement

```{r}
not_priority_binary(clean, goal_academics, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected improving academic achievement as a long-term goal", n_tab)
```

## By goal: Eliminate disparities in outcomes

```{r}
not_priority_binary(clean, goal_eliminate_disparities, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected eliminating disparities in student outcomes as a long-term goal", n_tab)
```

## By goal: Develop students' durable skills

```{r}
not_priority_binary(clean, goal_durable_skills, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected developing students' durable skills as a long-term goal", n_tab)
```

## By goal: Expand access to postsecondary education

```{r}
not_priority_binary(clean, goal_postsecondary, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected expanding access to postsecondary education as a long-term goal", n_tab)
```

## By goal: Strengthen career preparation

```{r}
not_priority_binary(clean, goal_career, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected strengthening career preparation as a long-term goal", n_tab)
```

## By goal: Improve instructional quality

```{r}
not_priority_binary(clean, goal_instruction, "Distribution of superintendents who did not indicate a goal needed substantial change, by whether they selected improving instructional quality as a long-term goal", n_tab)
```

## By belief: I believe my leadership positively influences outcomes for students in my district today.

```{r}
not_priority_cat(
  clean,
  op_positive_now,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: I believe my leadership positively influences outcomes for students in my district today.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: I believe my leadership will positively influence outcomes for students in my district even after I retire or move on.

```{r}
not_priority_cat(
  clean,
  op_positive_later,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: I believe my leadership will positively influence outcomes for students in my district even after I retire or move on.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: Addressing educational inequities is an essential part of my leadership responsibilities.

```{r}
not_priority_cat(
  clean,
  op_inequities,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: Addressing educational inequities is an essential part of my leadership responsibilities.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: I feel enough job security to take strategic risks.

```{r}
not_priority_cat(
  clean,
  op_job_security,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: I feel enough job security to take strategic risks.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: I notice a strong desire among district/central office staff to take on new challenges.

```{r}
not_priority_cat(
  clean,
  op_district_support,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: I notice a strong desire among district/central office staff to take on new challenges.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: I feel strong expectations from stakeholders to lead transformational change in my district.

```{r}
not_priority_cat(
  clean,
  op_expectations,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: I feel strong expectations from stakeholders to lead transformational change in my district.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: I feel personally motivated to lead transformational change in my district.

```{r}
not_priority_cat(
  clean,
  op_personal_motivation,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: I feel personally motivated to lead transformational change in my district.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: I feel confident in my ability to lead transformational change in my current district.

```{r}
not_priority_cat(
  clean,
  op_leadership_ability,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: I feel confident in my ability to lead transformational change in my current district.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: My formal education prepared me to lead change efforts.

```{r}
not_priority_cat(
  clean,
  op_preparation,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: My formal education prepared me to lead change efforts.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: In the past, I have been part of successful change efforts in a public school or district.

```{r}
not_priority_cat(
  clean,
  op_prior_success,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: In the past, I have been part of successful change efforts in a public school or district.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By belief: In the past, I have been part of stalled or failed change efforts in a public school or district.

```{r}
not_priority_cat(
  clean,
  op_prior_failure,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by response to: In the past, I have been part of stalled or failed change efforts in a public school or district.",
  recode_map = NULL,
  var_levels = c("Definitely true", "Mostly true", "Mostly false", "Definitely false")
)
```

## By state

```{r}
not_priority_cat(
  clean,
  state,
  n_tab,
  title = "Distribution of superintendents who did not indicate a goal needed substantial change, by state",
  recode_map = NULL,
  var_levels = c("CT", "MA", "ME", "NH", "RI", "VT")
)
```

```{r}
# Clean up
rm(not_priority_binary, not_priority_cat, name_map, desc)
```

# Percentage-point difference

Because some groups are larger than others in the overall sample, raw percentages like in the prior section can be misleading. To account for baseline group size, I looked at percentage-point differences between each group’s share among those who said some goal did not need significant change and its share in the full sample.  

```{r}
# set priority vars
priority <- clean %>% 
  select(starts_with("need") & !ends_with("other")) %>% 
  colnames()
# set grouping vars
group <- clean %>% 
  select(!c(starts_with("need"), ends_with("other"), id, role, partial_response)) %>% 
  colnames()
```

```{r}
# pull in function
source(here("functions/build_index_df.R"))
```

## Co-occurence with other responses

```{r}
# Clean up remaining cat variables
expanded <- clean %>% 
  mutate(across(starts_with("help"), ~case_when(
    . == 1 ~ "Definitely helps",
    . == 2 ~ "Helps more than hinders",
    . == 3 ~ "Hinders more than helps",
    . == 4 ~ "Definitely hinders"
  ))) %>% 
  mutate(across(starts_with("op"), ~case_when(
    . == 1 ~ "Definitely true",
    . == 2 ~ "Mostly true",
    . == 3 ~ "Mostly false",
    . == 4 ~ "Definitely false"
  ))) 
```


```{r}
# Create index df
index_df <- build_index_df(
  data = expanded,
  priority_cols = priority,
  group_vars = group,
  min_n_non = 5
)
```

Cells show the percentage-point difference between each group’s share among non-selectors and its share in the full sample. For example, a value of +3.4 means the group is 3.4 percentage points more common among non-selectors than in the overall sample. Positive values indicate overrepresentation; negative values indicate underrepresentation. Green cells mark values that were in the top percentile.  

[e.g., diff_pp= (% of people who did NOT select “safe, welcoming” and who DID select “strengthen career prep”) − (% of the full sample who selected “strengthen career prep”)]

```{r}
# Create display table
index_tab <- index_df %>% 
  select(group_var, group_level, priority, diff_pp) %>%
  left_join(labs, by = c("group_var" = "var_name")) %>% 
  mutate(
    group = nice_label,
    group_level = case_when(
      group_level == "1" ~ "Selected",
      group_level == "0" ~ "Did not select",
      TRUE ~ group_level
    ),
    row_label = paste(group, group_level, sep = ": ")
  ) %>%
  left_join(n_tab, by = c("priority" = "var_name")) %>% 
  left_join(labs, by = c("priority" = "var_name"), suffix = c("", "_priority")) %>%
  transmute(
    row_label,
    col_label = paste0(nice_label_priority, " (N = ", n, ")"),
    diff_pp
  ) %>%
  group_by(row_label, col_label) %>%
  summarize(diff_pp = mean(diff_pp, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = col_label, values_from = diff_pp)



# Identify numeric columns (all diff_pp columns)
num_cols <- names(index_tab)[sapply(index_tab, is.numeric)]

# Get 99th percentile across all numeric values
top_cutoff <- quantile(
  unlist(index_tab[num_cols]),
  probs = 0.99,
  na.rm = TRUE
)
```

(This is the full table for reference but it can be hard to read due to size so below I've included a version highlighting where there were notable differences)

```{r}
# Display table
datatable(
  index_tab,
  rownames = FALSE,
  options = list(
    pageLength = 25,
    scrollX = TRUE
  )
) %>%
  formatRound(columns = num_cols, digits = 1) %>%
  formatStyle(
    columns = num_cols,
    backgroundColor = styleInterval(
      top_cutoff,
      c("transparent", "#A5CE41")
    )
  )
```

This version zooms in on rows and goals where there were notable differences:

```{r}
# Identify numeric columns (all diff_pp columns)
num_cols <- names(index_tab)[sapply(index_tab, is.numeric)]

# Cutoff for "green" (you used 0.99 earlier; change to 0.90 for top 10%)
top_cutoff <- quantile(
  unlist(index_tab[num_cols]),
  probs = 0.99,
  na.rm = TRUE
)

# ---- Keep only rows that have at least one cell above the cutoff ----
index_tab_hi <- index_tab %>%
  rowwise() %>%
  mutate(has_hi = any(c_across(all_of(num_cols)) > top_cutoff, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(has_hi) %>%
  select(-has_hi)

# ---- Keep only columns that have at least one "green" value (within the filtered rows) ----
num_cols_hi <- names(index_tab_hi)[sapply(index_tab_hi, is.numeric)]

cols_with_hi <- num_cols_hi[
  sapply(index_tab_hi[num_cols_hi], function(x) any(x > top_cutoff, na.rm = TRUE))
]

index_tab_hi <- index_tab_hi %>%
  select(row_label, all_of(cols_with_hi))

# ---- Display table ----
datatable(
  index_tab_hi,
  rownames = FALSE,
  options = list(
    pageLength = 25,
    scrollX = TRUE
  )
) %>%
  formatRound(columns = cols_with_hi, digits = 1) %>%
  formatStyle(
    columns = cols_with_hi,
    backgroundColor = styleInterval(
      top_cutoff,
      c("transparent", "#A5CE41")
    )
  )
```


```{r}
# Clean up
rm(expanded, index_df, index_tab, num_cols, top_10_cutoff, index_tab_hi, cols_with_hi, group, num_cols_hi, top_cutoff)
```

## Demographic patterns

```{r}
# bring in demographic data
merged_dat <- import('/Users/janetteavelar/Library/CloudStorage/GoogleDrive-javelar3@asu.edu/.shortcut-targets-by-id/1NdfrCmhAW0NpUV47ruVeKThD8M3arhRq/DATA: Barr 3.0/Survey/Analysis/Current Data Files/MergedSurvey_251215.dta')
```

```{r}
# create vector of everything already cleaned except id
subset <- labs %>% filter(var != "qcode") %>% pull(var)
# clean
clean_merged <- merged_dat %>% 
  # strip labels
  haven::zap_label() %>% 
  # recode all missing values to NA
  mutate(across(where(is.numeric), ~na_if(., -99))) %>% 
  # drop already cleaned cols to merge back in
  select(!c(any_of(subset), grades_served, num_schools, num_high_schools,)) %>% 
  # back to cat...
  mutate(district_type = case_when(
      district_type == 1 ~ "Traditional",
      district_type == 2 ~ "Regional",
      district_type == 3 ~ "Other", #Regional CTE
      district_type == 4 ~ "Charter",
      district_type == 5 ~ "Other" #State agency
      ),
    urbanicity = case_when(
      urbanicity == 1 ~ "Urban",
      urbanicity == 2 ~ "Suburban",
      urbanicity == 3 ~ "Rural"),
    change_group = case_when(
      change_group == 1 ~ "0 domains",
      change_group == 2 ~ "1-2 domains",
      change_group == 3 ~ "3-5 domains",
      change_group == 4 ~ "6-9 domains"),
    enrollment_cat = ntile(enrollment_2025, 3),
    enrollment_cat = case_when(
      enrollment_cat == 1 ~ "Low",
      enrollment_cat == 2 ~ "Average",
      enrollment_cat == 3 ~ "High"
    ),
    enrollment_change_cat = ntile(enrollment_change_6yr_pct, 3),
    enrollment_change_cat = case_when(
      enrollment_change_cat == 1 ~ "Low",
      enrollment_change_cat == 2 ~ "Average",
      enrollment_change_cat == 3 ~ "High"
    ),
    pct_bipoc = 1 - pct_white,
    pct_bipoc_cat = ntile(pct_bipoc, 3),
    pct_bipoc_cat = case_when(
      pct_bipoc_cat == 1 ~ "Low",
      pct_bipoc_cat == 2 ~ "Average",
      pct_bipoc_cat == 3 ~ "High"
    ),
    pct_lowinc_cat = ntile(pct_lowinc, 3),
    pct_lowinc_cat = case_when(
      pct_lowinc_cat == 1 ~ "Low",
      pct_lowinc_cat == 2 ~ "Average",
      pct_lowinc_cat == 3 ~ "High"
    ),
    pct_el_cat = ntile(pct_el, 3),
    pct_el_cat = case_when(
      pct_el_cat == 1 ~ "Low",
      pct_el_cat == 2 ~ "Average",
      pct_el_cat == 3 ~ "High"
    ),
    pct_sped_cat = ntile(pct_sped, 3),
    pct_sped_cat = case_when(
      pct_sped_cat == 1 ~ "Low",
      pct_sped_cat == 2 ~ "Average",
      pct_sped_cat == 3 ~ "High"
    ),
    pct_math_cat = ntile(pct_math_prof_3_8, 3),
    pct_math_cat = case_when(
      pct_math_cat == 1 ~ "Low",
      pct_math_cat == 2 ~ "Average",
      pct_math_cat == 3 ~ "High"
    ),
    pct_ela_cat = ntile(pct_ela_prof_3_8, 3),
    pct_math_cat = case_when(
      pct_ela_cat == 1 ~ "Low",
      pct_ela_cat == 2 ~ "Average",
      pct_ela_cat == 3 ~ "High"
    ),
    pct_grad_cat = ntile(grad_rate, 3),
    pct_grad_cat = case_when(
      pct_grad_cat == 1 ~ "Low",
      pct_grad_cat == 2 ~ "Average",
      pct_grad_cat == 3 ~ "High"
    ),
    serve_hs = ifelse(serve_hs == 1, "Serves high school", "Does not serve high school"))
# set grouping vars you're interested in
group <- clean_merged %>% select(district_type, urbanicity, serve_hs, change_group, ends_with("cat")) %>% colnames()

#merge with main dataset
clean_merged <- clean_merged %>% 
  right_join(clean, by = c("qcode" = "id"))
```

```{r}
# Create index df
index_df <- build_index_df(
  data = clean_merged,
  priority_cols = priority,
  group_vars = group,
  min_n_non = 5
)
```

```{r}
# Create display table
index_tab <- index_df %>% 
  select(group_var, group_level, priority, diff_pp) %>%
  mutate(row_label = paste(group_var, group_level, sep = ": ")) %>%
  left_join(n_tab, by = c("priority" = "var_name")) %>% 
  left_join(labs, by = c("priority" = "var_name"), suffix = c("", "_priority")) %>%
  transmute(
    row_label,
    col_label = paste0(nice_label, " (N = ", n, ")"),
    diff_pp
  ) %>%
  group_by(row_label, col_label) %>%
  summarize(diff_pp = mean(diff_pp, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = col_label, values_from = diff_pp)



# Identify numeric columns (all diff_pp columns)
num_cols <- names(index_tab)[sapply(index_tab, is.numeric)]

# Upper tail (top 1%)
upper_cutoff <- quantile(
  unlist(index_tab[num_cols]),
  probs = 0.99,
  na.rm = TRUE
)

# Lower tail (bottom 1%)
lower_cutoff <- quantile(
  unlist(index_tab[num_cols]),
  probs = 0.01,
  na.rm = TRUE
)
```

(This is the full table for reference but it can be hard to read due to size so below I've included a version highlighting where there were notable differences)

```{r}
# Display table
datatable(index_tab, rownames = FALSE) %>%
  formatRound(columns = num_cols, digits = 1) %>%
  formatStyle(
    columns = num_cols,
    backgroundColor = styleInterval(
      c(lower_cutoff, upper_cutoff),
      c("#D73027", "transparent", "#A5CE41")  # red for low, green for high
    )
  )
```

This version zooms in on rows and goals where there were notable differences:

```{r}
# ---- Keep only rows that have at least one cell above the cutoff ----
index_tab_hi <- index_tab %>%
  rowwise() %>%
  mutate(
    has_extreme = any(
      c_across(all_of(num_cols)) > upper_cutoff |
      c_across(all_of(num_cols)) < lower_cutoff,
      na.rm = TRUE
    )
  ) %>%
  ungroup() %>%
  filter(has_extreme) %>%
  select(-has_extreme)

# Identify numeric columns after row filtering
num_cols_hi <- names(index_tab_hi)[sapply(index_tab_hi, is.numeric)]

# Keep columns that contain at least one extreme (upper OR lower tail)
cols_with_hi <- num_cols_hi[
  sapply(index_tab_hi[num_cols_hi], function(x)
    any(x > upper_cutoff | x < lower_cutoff, na.rm = TRUE)
  )
]

# Keep only row_label + extreme columns
index_tab_hi <- index_tab_hi %>%
  select(row_label, all_of(cols_with_hi))

# ---- Display table ----
datatable(
  index_tab_hi,
  rownames = FALSE,
  options = list(pageLength = 25, scrollX = TRUE)
) %>%
  formatRound(columns = cols_with_hi, digits = 1) %>%
  formatStyle(
    columns = cols_with_hi,
    backgroundColor = styleInterval(
      c(lower_cutoff, upper_cutoff),
      c("#D73027", "transparent", "#A5CE41")  # red for low, green for high
    )
  )
```

# TLDR

**Response patterns**
Most subgroup differences in goals that were "deprioritized" (i.e., indicated as *not* needing substantial change) were modest, typically within ±1–2 percentage points. However, some cells fell in the top 1% of observed differences (marked in green). Surprisingly, these occurred within priority areas that had more respondents (N somewhere between ~25 and ~90), suggesting that they are not solely artifacts of small sample size.  

The largest differences did not cluster primarily around tenure or structural demographic characteristics. Instead, they were more closely associated with leaders’ stated goals. In particular, superintendents who selected “Increase student engagement” as a priority were more likely than the overall sample to think certain goals did not need significant change, including:  
-how adults treat students
-budget alignment to strategic goals
-course placement practices  

While the magnitude of these differences was moderate (generally in the 3–6 percentage-point range), their recurrence across multiple priorities suggests a slight pattern rather than isolated variation.  

There was no evidence of sharp divides along tenure or other structural categories. The outliers seem to reflect variation in goals and strategies rather than clear polarization across leader characteristics, as far as we can observe them from other survey questions.  

**Demographics**
When demographic and district characteristics are incorporated, clearer structural patterns emerge and there are more pronounced subgroup differences. Larger deviations are associated with district size, enrollment changes, whether or not the district has a high school, and overall breadth of perceived change needs (i.e., change group).  

Low-enrollment and growing districts are more likely to view areas such as course placement, budget alignment, and how adults treat students as requiring substantial change. In contrast, districts that serve high school grades and those in one of the high change groups (3-5) are more likely to indicate that course placement, budget alignment, and how adults treat students do not need substantial change.  

These patterns suggest that perceptions of need are shaped by district context, specifically related to size, growth, and grade spans which appear to influence how leaders assess the urgency of these goals.  