---
title: "Modeling Change Efforts"
author: "Janette Avelar"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: flatly
    highlight: tango
    code_folding: hide
  word_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # hides code
  message = FALSE, # hides messages
  warning = FALSE,  # hides warnings
  fig.width = 10, fig.height = 9
)
library(rio)
library(tidyverse)
library(here)
library(DT)
library(kableExtra)
library(lme4)
library(modelsummary)
library(fixest)
library(car)
library(stringr)
library(marginaleffects)
#library(MASS) #note: masks select()

dat <- import('/Users/janetteavelar/Library/CloudStorage/GoogleDrive-javelar3@asu.edu/.shortcut-targets-by-id/1NdfrCmhAW0NpUV47ruVeKThD8M3arhRq/DATA: Barr 3.0/Survey/Analysis/Current Data Files/MergedSurvey_251215.dta')
```

Instructions: `...look at the analysis but using alternative outcome variables - continuous variable and a dichotomous variable that is more change (3+) vs less change (0-2). same set of variable to test against (all the variables you tested as well as the goals (q3), the change needed (Q4+Q5), and the will/motivation (q11 + q12).`

```{r}
# Clean up data
clean <- dat %>% 
  # remove labels
  haven::zap_label() %>% 
  # handle missing values
  mutate(
  across(where(is.numeric), ~ na_if(., -99)),
  across(where(is.character), ~ na_if(., "-99"))) %>% 
  # change numeric -> categorical
  mutate(grades_served = case_when(
    grades_served == 1 ~ "K-6",
    grades_served == 2 ~ "K-8",
    grades_served == 3 ~ "K-12",
    grades_served == 4 ~ "6-8",
    grades_served == 5 ~ "7-12",
    grades_served == 6 ~ "9-12"
  ),
  district_type = case_when(
    district_type == 1 ~ "Traditional",
    district_type == 2 ~ "Regional",
    district_type == 3 ~ "Regional CTE",
    district_type == 4 ~ "Charter",
    district_type == 5 ~ "State agency"
  ),
  urbanicity = case_when(
    urbanicity == 1 ~ "Urban",
    urbanicity == 2 ~ "Suburban",
    urbanicity == 3 ~ "Rural"
  ),
  ma_account_code = case_when(
    ma_account_code == 1 ~ "Insufficient data",
    ma_account_code == 2 ~ "Not requiring assistance",
    ma_account_code == 3 ~ "Requiring assistance or intervention"
  ),
  change_group_dich = case_when(
    change_group == 1 | change_group == 2 ~ 0,
    change_group == 3 | change_group == 4 ~ 1
  ),
  serve_hs = case_when(
    serve_hs == 0 ~ "Does not serve high school",
    serve_hs == 1 ~ "Serves high school"
  )) %>% 
  # build numeric outcome
  filter(!is.na(change_group)) %>% 
  group_by(qcode) %>% 
  mutate(n_efforts = rowSums(across(starts_with("q6")), na.rm = TRUE)) %>% 
  ungroup()
```

# V1

## Continuous

Our continuous specification uses original survey responses for Q6, which indicate up to 9 potential areas for change efforts that a district may be taking on, selected from 17 possible responses.  

Below are the descriptive statistics for the continuous change efforts variable. Initially I planned to run a Poisson model, but because the variance is so much larger than the mean I use a negative binomial regression model below instead. These type of models are designed for count outcomes when there is a lot of variation across cases and may fit the data better than a simpler approach. In other words, the model accounts for the fact that some schools have consistently higher or lower numbers of efforts than others, rather than forcing all schools to follow the same pattern.  

The results of these models are the log of the predicted count, which I have exponentiated so we can interpret them as percent changes in the expected number of efforts, holding other factors constant. In this type of model, 1 is our baseline of "no change" rather than 0. For example, a value of 1.3 would indicate a 30% increase, while a value of .6 would indicate a 40% decrease.  

```{r}
desc <- psych::describe(clean$n_efforts) %>% as_tibble()
variance <- round(var(clean$n_efforts), 3)
desc %>% 
  mutate(
    variable = "N change efforts",
    variance = variance) %>% 
  select(variable, n, mean, sd, min, max, se, variance) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  DT::datatable()
```

I'm fitting multiple models with covariate blocks that include...  

Structural/contextual covariates:  
-state  
-urbanicity  
-district type (disrict/charter/other)  
-size (centered enrollment)  
-grades (elementary-only, secondary-only, elementary+secondary)  

Student composition:  
-centered % black/latine  
-centered % low-income  
-centered % EL  
-centered % SPED  

Academic performance  
-centered % proficient 3-8 (dropped HS because only grade 11 tested)  
-MAYBE centered % graduation rate but will drop elementary-only  

Leadership goals and beliefs:  
-breadth of goals (avg. across increase student engagement, safe/welcoming community, academic achievement, eliminate disparities, durable skills, postsecondary access, career prep, instructional quality)  
-perceived need for instructional/relational change (avg. across what teachers teach, how teachers teach, assessing student learning, addressing non-academic needs, how adults treat students, what adults think students are capable of, student decision-making, course placement)  
-perceived need for systemic and structural change (avg. across supporting teachers, supporting principals, staffing organization, budget alignment, data use, family engagement, AI use)  
-will and confidence (avg. across leadership influences outcomes now + long-term, addressing inequities is essential, job security to take risks, staff desire for challenges, expectations to lead change, personal motivation, confidence leading change, prior successful change experience, prior stalled or failed change efforts reverse-coded)  

```{r}
# additional covariate cleaning
mod <- clean %>% 
  # structural/contextual
  mutate(district_type = ifelse((district_type == "Regional" | district_type == "Regional CTE" | district_type == "CTE" | district_type == "State agency"), "Other type", district_type),
         c_enrollment = as.numeric(scale(enrollment_2025)),
         c_enrollment_change = as.numeric(scale(enrollment_change_6yr_pct)),
         grades_served = case_when(
           grades_served == "6-8" ~ "Secondary only",
           grades_served == "7-12" ~ "Secondary only",
           grades_served == "9-12" ~ "Secondary only",
           grades_served == "K-12" ~ "Elementary + Secondary",
           grades_served == "K-6" ~ "Elementary only",
           grades_served == "K-8" ~ "Elementary + Secondary"
         )) %>% 
  # student composition
  mutate(pct_blacklatine = pct_black + pct_hispanic,
         c_pct_blacklatine = as.numeric(scale(pct_blacklatine)),
         c_pct_lowincome = as.numeric(scale(pct_lowinc)),
         c_pct_el = as.numeric(scale(pct_el)),
         c_pct_sped = as.numeric(scale(pct_sped))) %>% 
  # academic performance
  rename(c_ela_3_8 = pct_ela_prof_3_8_center,
         c_math_3_8 = pct_math_prof_3_8_center,
         c_grad = grad_rate_center) %>%
  # leadership: breadth of goals
  mutate(breadth_goals = rowMeans(across(q3_1:q3_9), na.rm = TRUE),
         c_breadth_goals = as.numeric(scale(breadth_goals))) %>% 
  # leadership: need for instructional/relational change
  mutate(instructional_change = rowMeans(across(q4_1:q4_8), na.rm = TRUE),
         c_instructional_change = as.numeric(scale(instructional_change))) %>% 
  # leadership: need for systemic/structural change
  mutate(systemic_change = rowMeans(across(q5_1:q5_7), na.rm = TRUE),
         c_systemic_change = as.numeric(scale(systemic_change))) %>% 
  # leadership: will and confidence
  mutate(q12_6 = case_when(
    q12_6 == 1 ~ 4,
    q12_6 == 2 ~ 3,
    q12_6 == 3 ~ 2,
    q12_6 == 4 ~ 1),
    will_confidence = rowMeans(across(q11_1:q12_6), na.rm= TRUE),
    c_will_confidence = as.numeric(scale(will_confidence))) 
```

The table below presents the modeling results tests the association between the number of efforts a district is working on with each block of covariates one at a time. Note that the composite for systemic change and instructional change throws a warning that there is no sign of overdispersion and we may consider a Poisson model. We can ignore that warning for now.  

Some noticings from the table:  
-traditional and "other" districts have a higher number of change efforts relative to charter schools, and this was marginally sig. for "other"  
-elementary schools have fewer change efforts relative to combined elementary/secondary districts (sig.)  
-student demographics and enrollment don't seem to have much of an effect on the predicted number of change efforts (values near 1)  
-ELA achievement similarly does not substantially change predicted number of change efforts, but a standard deviation increase in math proficiency does - by a reduction of 50%, i.e., districts that are doing well in math are not adopting as many efforts  
-leader confidence and breadth of goals don't shift the number of change efforts very much - I was very surprised by this, particularly for the latter which I expected would have a positive relationship, but it may be because these are composites of all items and we could consider paring down to a composite of only those we think are most relevant

```{r}
# Build models
#m0 <- glm.nb(n_efforts ~ 1, data = mod) # MASS version
m0 <- fenegbin(n_efforts ~ 1, data = mod)
m1 <- update(m0, . ~ district_type + c_enrollment + c_enrollment_change +  grades_served)
m2 <- update(m0, . ~ c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m0, . ~ c_ela_3_8 + c_math_3_8 + c_grad)
m4 <- update(m0, . ~ c_breadth_goals + c_instructional_change + c_systemic_change + c_will_confidence)
```

```{r}
# Display results function
stepwise_table <- function(m0 = m0, m1 = m1, m2 = m2, m3 = m3, m4 = m4){
  table <- modelsummary(list(
  "Null" = m0,
  "Structural/ Contextual" = m1,
  "Student composition" = m2,
  "Academic performance" = m3,
  "Leader attitude" = m4
),
exponentiate = TRUE,
coef_map = c(
  "(Intercept)" = "Intercept",
  "district_typeOther type" = "Other type",
  "district_typeTraditional" = "Traditional",
  "district_typeCharter" = "Charter",
  "c_enrollment" = "Enrollment",
  "c_enrollment_change" = "6-year enrollment change",
  "grades_servedElementary only" = "Elementary only",
  "grades_servedSecondary only" = "Secondary only",
  "grades_servedElementary + Secondary" = "Elementary + Secondary",
  "c_pct_blacklatine" = "% Black/Latine",
  "c_pct_lowincome" = "% Low income",
  "c_pct_el" = "% EL",
  "c_pct_sped" = "% SPED",
  "c_ela_3_8" = "% ELA Proficient",
  "c_math_3_8" = "% Math Proficient",
  "c_grad" = "% Graduating in 4 years",
  "c_breadth_goals" = "Breadth of goals",
  "c_instructional_change" = "Perceived need for instructional/relational change",
  "c_systemic_change" = "Perceived need for systemic change",
  "c_will_confidence" = "Leader will and confidence"
),
coef_omit = "theta",
stars = TRUE,
estimate = "{estimate}{stars} ({std.error})",
statistic = NULL,
fmt = 2)
  
  return(table)
}
```

```{r}
stepwise_table(m0, m1, m2, m3, m4)
```

Next, I've integrated the composites into a single model in the table below, again using a stepwise approach. For this model, to be able to compare model fit, I've limited the sample to the 204 districts with complete data on all covariates. With this model, the only statistically significant predictors are a perceived need for instructional/relational change and systemic change. Most other predictors are hovering around 1, with the exception of achievement - more students doing well in ELA translate to more change efforts while the opposite is true for math. Higher graduation rates also translate to fewer change efforts.

```{r}
# Ensure sample is the same
vars <- c(
  "n_efforts",
  "district_type", "c_enrollment", "grades_served",
  "c_pct_blacklatine", "c_pct_lowincome", "c_pct_el", "c_pct_sped",
  "c_ela_3_8", "c_math_3_8", "c_grad",
  "c_breadth_goals", "c_instructional_change", "c_systemic_change", "c_will_confidence"
)
mod_cc <- mod %>%
  select(all_of(vars)) %>%
  filter(complete.cases(.))
```

```{r}
# Build stepwise models
m0 <- fenegbin(n_efforts ~ 1, data = mod_cc)
m1 <- update(m0, . ~ . + district_type + c_enrollment + grades_served)
m2 <- update(m1, . ~ . + c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m2, . ~ . + c_ela_3_8 + c_math_3_8 + c_grad)
m4 <- update(m3, . ~ . + c_breadth_goals + c_instructional_change + c_systemic_change + c_will_confidence)
```

```{r}
# display table
stepwise_table(m0, m1, m2, m3, m4)
```

## Dichotomous

Our dichotomous specification uses original survey responses for Q6, which indicate up to 9 potential areas for change efforts that a district may be taking on, selected from 17 possible responses. We've split them into two categories: fewer change efforts (0-2) and more change efforts (3+).  

I use a logistic regression model to account for the binary outcome, with results presented in exponentiated form similarly to the negative binomial models above, but in odds ratios (rather than as a log of the expected count). In this type of model, a value of 1 indicates no difference in the odds. Values greater than 1 indicate higher odds of reporting high levels of change, while values less than 1 indicate lower odds. For example, a value of 2 indicates that the odds are twice as high, whereas a value of 0.5 indicates that the odds are half as high, holding other factors constant.    

The table below shows the descriptive statistics for this specification of the outcome variable.  

```{r}
desc <- psych::describe(clean$change_group_dich) %>% as_tibble()
variance <- round(var(clean$change_group_dich), 3)
desc %>% 
  mutate(
    variable = "N change efforts",
    variance = variance) %>% 
  select(variable, n, mean, sd, min, max, se, variance) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  DT::datatable()
```

The table below presents the modeling results tests the association between the number of efforts a district is working on with each block of covariates one at a time.   

Some noticings from the table:  
-traditional and "other" districts have twice the odds of adopting 3 or more change efforts compared to a charter district (n.s.)  
-districts that only serve elementary or secondary grades have lower odds relative to those that serve across the K-12 span (n.s.)  
-with the binary specification, we see the relationship between achievement and change efforts flip (compared to the negative binomial model) - now, higher ELA proficiency is associated with lower odds, and higher math proficiency doesn't change the odds (n.s.)  
-higher graduation rates lower the odds of adopting 3+ change efforts (n.s.)  
-like in our negative binomial models, a perceived need for instructional/relational change and a perceived need for systemic change are the only statistically significant predictors of a greater number of change efforts  

```{r}
# Build models
m0 <- feglm(change_group_dich ~ 1, data = mod, family = binomial())
m1 <- update(m0, . ~ district_type + c_enrollment + c_enrollment_change +  grades_served)
m2 <- update(m0, . ~ c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m0, . ~ c_ela_3_8 + c_math_3_8 + c_grad)
m4 <- update(m0, . ~ c_breadth_goals + c_instructional_change + c_systemic_change + c_will_confidence)
```

```{r}
stepwise_table(m0, m1, m2, m3, m4)
```

Next, I've integrated the composites into a single model in the table below, again using a stepwise approach. For this model, to be able to compare model fit, I've limited the sample to the 204 districts with complete data on all covariates. The fully specified model had the strongest fit.  

With this model, the only statistically significant predictors are a perceived need for instructional/relational change and systemic change, and are each positively associated with the odds of undertaking more change efforts. Most other predictors are hovering around even odds, with the exception of achievement - more students doing well in ELA translates to more change efforts while the opposite is true for math. Higher graduation rates also translate to fewer change efforts.  

```{r}
# Build stepwise models
m0 <- fenegbin(n_efforts ~ 1, data = mod_cc)
m1 <- update(m0, . ~ . + district_type + c_enrollment + grades_served)
m2 <- update(m1, . ~ . + c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m2, . ~ . + c_ela_3_8 + c_math_3_8 + c_grad)
m4 <- update(m3, . ~ . + c_breadth_goals + c_instructional_change + c_systemic_change + c_will_confidence)
```

```{r}
# display table
stepwise_table(m0, m1, m2, m3, m4)
```

```{r}
#Clean up
rm(m0, m1, m2, m3, m4)
```


# V2

## Continuous

Our continuous specification uses original survey responses for Q6, which indicate up to 9 potential areas for change efforts that a district may be taking on, selected from 17 possible responses.  

Like in V1, I'm using a negative binomial model for the continuous outcome to account for it being a count variable with variance larger than the mean. The results of these models are the log of the predicted count, which I have exponentiated so we can interpret them as percent changes in the expected number of efforts, holding other factors constant. In this type of model, 1 is our baseline of "no change" rather than 0. For example, a value of 1.3 would indicate a 30% increase, while a value of .6 would indicate a 40% decrease.  

I'm fitting multiple models with covariate blocks that include...  

Structural/contextual covariates:  
-urbanicity  
-district type (disrict/charter/other)  
-size (centered enrollment)  
-presence of high school  

Student composition:  
-centered % black/latine  
-centered % low-income  
-centered % EL  
-centered % SPED  

Academic performance  
-centered % proficient 3-8 (dropped HS because only grade 11 tested)  

Leadership goals and beliefs:  
-length of tenure  
-whether student engagement was selected as a goal  
-whether instructional quality was selected as a goal  
-leader confidence that leadership influences outcomes now  
-reported personal motivation  
-reported confidence leading change in the district  

**IMPORTANT NOTE:** I used q2 to construct length of tenure but it differs from the codebook. The codebook specifies there should be **two** q2 variables: q2a and q2b. What is q2??  

And state fixed effects once the model is fully specified.  

First, I use a stepwise approach to check how the covariates in each block behave.  

**Structural/contextual**

```{r}
# Build models
m0 <- fenegbin(n_efforts ~ 1, data = mod)
m1 <- update(m0, . ~ . + urbanicity)
m2 <- update(m0, . ~ . + district_type)
m3 <- update(m0, . ~ . + c_enrollment)
m4 <- update(m0, . ~ . + serve_hs)
```

```{r}
modelsummary(list(m0, m1, m2, m3, m4),
             exponentiate = TRUE,
             coef_map = c(
               "(Intercept)" = "Intercept",
               "district_typeOther type" = "Other type",
               "district_typeTraditional" = "Traditional",
               "district_typeCharter" = "Charter",
               "c_enrollment" = "Enrollment",
               "urbanicitySuburban" = "Suburban",
               "urbanicityUrban" = "Urban",
               "urbanicityRural" = "Rural"),
             coef_omit = "theta",
             gof_omit = "BIC|Log\\.Lik\\.|R2|Adj\\.|Pseudo|RMSE",
             stars = TRUE,
             estimate = "{estimate}{stars} ({std.error})",
             statistic = NULL,
             fmt = 2)
```

```{r}
rm(m0, m1, m2, m3, m4)
```

**Student composition**

```{r}
# Build models
m0 <- fenegbin(n_efforts ~ 1, data = mod)
m1 <- update(m0, . ~ . + c_pct_blacklatine)
m2 <- update(m0, . ~ . + c_pct_lowincome)
m3 <- update(m0, . ~ . + c_pct_el)
m4 <- update(m0, . ~ . + c_pct_sped)
```

```{r}
modelsummary(list(m0, m1, m2, m3, m4),
             exponentiate = TRUE,
             coef_map = c(
               "(Intercept)" = "Intercept",
               "c_pct_blacklatine" = "% Black/Latine",
               "c_pct_lowincome" = "% Low income",
               "c_pct_el" = "% EL",
               "c_pct_sped" = "% SPED"),
             coef_omit = "theta",
             gof_omit = "BIC|Log\\.Lik\\.|R2|Adj\\.|Pseudo|RMSE",
             stars = TRUE,
             estimate = "{estimate}{stars} ({std.error})",
             statistic = NULL,
             fmt = 2)
```

```{r}
rm(m0, m1, m2, m3, m4)
```

**Academic performance**  

```{r}
# Build models
m0 <- fenegbin(n_efforts ~ 1, data = mod)
m1 <- update(m0, . ~ . + c_ela_3_8)
m2 <- update(m1, . ~ . + c_math_3_8)
```

```{r}
modelsummary(list(m0, m1, m2),
             exponentiate = TRUE,
             coef_map = c(
               "(Intercept)" = "Intercept",
               "c_ela_3_8" = "% ELA Proficient",
               "c_math_3_8" = "% Math Proficient"),
             coef_omit = "theta",
             gof_omit = "BIC|Log\\.Lik\\.|R2|Adj\\.|Pseudo|RMSE",
             stars = TRUE,
             estimate = "{estimate}{stars} ({std.error})",
             statistic = NULL,
             fmt = 2)
```

```{r}
rm(m0, m1, m2)
```

**Leadership goals and beliefs**

```{r}
# add length of tenure
mod <- mod %>% 
  mutate(length_tenure = case_when(
    q2 == 1 ~ "Less than 1 year",
    q2 == 2 ~ "1–3 years",
    q2 == 3 ~ "4-6 years",
    q2 == 4 ~ "7–10 years",
    q2 == 5 ~ "More than 10 years"
  ),
  leadership_influence = case_when(
    q11_1 == 1 ~ 1,
    q11_1 == 2 ~ 1,
    q11_1 == 3 ~ 0,
    q11_1 == 4 ~ 0
  ),
  personal_motivation = case_when(
    q12_2 == 1 ~ 1,
    q12_2 == 2 ~ 1,
    q12_2 == 3 ~ 0,
    q12_2 == 4 ~ 0
  ),
  confidence_leading_change = case_when(
    q12_3 == 1 ~ 1,
    q12_3 == 2 ~ 1,
    q12_3 == 3 ~ 0,
    q12_3 == 4 ~ 0
  )) %>% 
  rename(student_engagement = q3_1,
         instructional_quality = q3_8)
```

```{r}
# Build models
m0 <- fenegbin(n_efforts ~ 1, data = mod)
m1 <- update(m0, . ~ . + length_tenure)
m2 <- update(m0, . ~ . + student_engagement)
m3 <- update(m0, . ~ . + instructional_quality)
m4 <- update(m0, . ~ . + leadership_influence)
m5 <- update(m0, . ~ . + personal_motivation)
m6 <- update(m0, . ~ . + confidence_leading_change)
```

```{r}
modelsummary(list(m0, m1, m2, m3, m4, m5, m6),
             exponentiate = TRUE,
             coef_map = c(
               "(Intercept)" = "Intercept",
               "confidence_leading_change" = "Confidence in ability to lead change",
               "personal_motivation" = "Personal motivation to lead change",
               "leadership_influence" = "Believes leadership positively influences outcomes now",
               "instructional_quality" = "Goal to improve academic achievement",
               "student_engagement" = "Goal to increase student engagement",
               "length_tenureLess than 1 year" = "Tenure less than 1 year",
               "length_tenure4-6 years" = "Tenure 4-6 years",
               "length_tenure7–10 years" = "Tenure 7–10 years",
               "length_tenureMore than 10 years" = "Tenure more than 10 years"),
             coef_omit = "theta",
             gof_omit = "BIC|Log\\.Lik\\.|R2|Adj\\.|Pseudo|RMSE",
             stars = TRUE,
             estimate = "{estimate}{stars} ({std.error})",
             statistic = NULL,
             fmt = 2)
```

```{r}
rm(m0, m1, m2, m3, m4, m5, m6)
```

Checking collinearity:

The only meaningful collinearity in the model is between ELA and math achievement; all other predictors exhibit low variance inflation. We could consider creating an index/composite measure if we want to reduce noise and be able to say something about acheivement, but otherwise should just be cautious with interpretation and use them only as controls/not for interpretation. We could consider the same for % low income and % black/latine, but these are within an acceptable level (<3).

```{r}
vif_mod <- lm(
  n_efforts ~ 
    district_type + c_enrollment + c_enrollment_change + grades_served +
    c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped +
    c_ela_3_8 + c_math_3_8 + c_grad +
    c_breadth_goals + c_instructional_change +
    c_systemic_change + c_will_confidence,
  data = mod
)
vif(vif_mod)
rm(vif_mod)
```


The table below presents the modeling results tests the association between the number of efforts a district is working on with each block of covariates one at a time.   

Some noticings from the table:  
-longer tenure is associated with slightly lower odds (n.s.)  
-goal to increase student engagement is associated with higher odds (sig) as is improving academic achievement (n.s.)  
-believing that leadership positively influences change reduced the odds by half (marginally sig.)  
-ELA proficiency associated with lower odds (n.s.) but note that math and ELA are highly correlated (.94) so we should be cautious of interpretation  
-% SPED has a very slight increase in odds (sig.)  

```{r}
# Build models
m0 <- fenegbin(n_efforts ~ 1, data = mod)
m1 <- update(m0, . ~ + urbanicity + district_type + c_enrollment + serve_hs)
m2 <- update(m0, . ~ + c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m0, . ~ + c_ela_3_8 + c_math_3_8)
m4 <- update(m0, . ~ + length_tenure + student_engagement + instructional_quality + leadership_influence + personal_motivation + confidence_leading_change)
```

```{r}
# Display results function
stepwise_table <- function(m0 = m0, m1 = m1, m2 = m2, m3 = m3, m4 = m4){
  table <- modelsummary(list(
  "Null" = m0,
  "Structural/ Contextual" = m1,
  "Student composition" = m2,
  "Academic performance" = m3,
  "Leader attitude" = m4
),
exponentiate = TRUE,
coef_map = c(
  "(Intercept)" = "Intercept",
  "district_typeOther type" = "Other type",
  "district_typeTraditional" = "Traditional",
  "district_typeCharter" = "Charter",
  "c_enrollment" = "Enrollment",
  "urbanicitySuburban" = "Suburban",
  "urbanicityUrban" = "Urban",
  "urbanicityRural" = "Rural",
  "c_pct_blacklatine" = "% Black/Latine",
  "c_pct_lowincome" = "% Low income",
  "c_pct_el" = "% EL",
  "c_pct_sped" = "% SPED",
  "c_ela_3_8" = "% ELA Proficient",
  "c_math_3_8" = "% Math Proficient",
  "confidence_leading_change" = "Confidence in ability to lead change",
  "personal_motivation" = "Personal motivation to lead change",
  "leadership_influence" = "Believes leadership positively influences outcomes now",
   "instructional_quality" = "Goal to improve academic achievement",
   "student_engagement" = "Goal to increase student engagement",
   "length_tenureLess than 1 year" = "Tenure less than 1 year",
   "length_tenure4-6 years" = "Tenure 4-6 years",
   "length_tenure7–10 years" = "Tenure 7–10 years",
   "length_tenureMore than 10 years" = "Tenure more than 10 years"
),
coef_omit = "theta",
gof_omit = "BIC|Log\\.Lik\\.|R2|Adj\\.|Pseudo|RMSE",
stars = TRUE,
estimate = "{estimate}{stars} ({std.error})",
statistic = NULL,
fmt = 2)
  
  return(table)
}
```

```{r}
stepwise_table(m0, m1, m2, m3, m4)
```

Next, I've integrated the composites into a single model in the table below, again using a stepwise approach. For this model, to be able to compare model fit, I've limited the sample to the 262 districts with complete data on all covariates.  

Noticings:  
-nearly all stat. sig. goes away  
-traditional/other districts have more change (n.s.)  
-urban have slightly fewer changes (n.s.)  
-goals slightly increase change (n.s./marginally sig.)  
-belief in leadership influencing outcomes decreases change (n.s.)  
-longer tenure is associated with fewer changes (only sig. at 10+ years)  

```{r}
# Ensure sample is the same
vars <- c(
  "n_efforts",
  "urbanicity", "district_type", "c_enrollment", "serve_hs",
  "c_pct_blacklatine", "c_pct_lowincome", "c_pct_el", "c_pct_sped",
  "c_ela_3_8", "c_math_3_8",
  "length_tenure", "student_engagement", "instructional_quality", "leadership_influence", "personal_motivation", "confidence_leading_change",
  "state"
)
mod_cc <- mod %>%
  select(all_of(vars)) %>%
  filter(complete.cases(.))
```

```{r}
# Build stepwise models
m0 <- fenegbin(n_efforts ~ 1, data = mod_cc)
m1 <- update(m0, . ~ . + urbanicity + district_type + c_enrollment + serve_hs)
m2 <- update(m1, . ~ . +  c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m2, . ~ . + c_ela_3_8 + c_math_3_8)
m4 <- update(m3, . ~ . + length_tenure + student_engagement + instructional_quality + leadership_influence + personal_motivation + confidence_leading_change)
```

```{r}
# display table
stepwise_table(m0, m1, m2, m3, m4)
```

And finally, I run the fully specified model once more with state fixed effects to control for variation across context.

```{r}
m5 <- update(m4, . ~ . | state)
```

```{r}
modelsummary(list(
  "Fully specified" = m4,
  "+ State FE" = m5
),
exponentiate = TRUE,
coef_map = c(
  "(Intercept)" = "Intercept",
  "district_typeOther type" = "Other type",
  "district_typeTraditional" = "Traditional",
  "district_typeCharter" = "Charter",
  "c_enrollment" = "Enrollment",
  "urbanicitySuburban" = "Suburban",
  "urbanicityUrban" = "Urban",
  "urbanicityRural" = "Rural",
  "c_pct_blacklatine" = "% Black/Latine",
  "c_pct_lowincome" = "% Low income",
  "c_pct_el" = "% EL",
  "c_pct_sped" = "% SPED",
  "c_ela_3_8" = "% ELA Proficient",
  "c_math_3_8" = "% Math Proficient",
  "confidence_leading_change" = "Confidence in ability to lead change",
  "personal_motivation" = "Personal motivation to lead change",
  "leadership_influence" = "Believes leadership positively influences outcomes now",
   "instructional_quality" = "Goal to improve academic achievement",
   "student_engagement" = "Goal to increase student engagement",
   "length_tenureLess than 1 year" = "Tenure less than 1 year",
   "length_tenure4-6 years" = "Tenure 4-6 years",
   "length_tenure7–10 years" = "Tenure 7–10 years",
   "length_tenureMore than 10 years" = "Tenure more than 10 years"
),
coef_omit = "theta",
gof_omit = "BIC|Log\\.Lik\\.|R2|Adj\\.|Pseudo|RMSE",
stars = TRUE,
estimate = "{estimate}{stars} ({std.error})",
statistic = NULL,
fmt = 2)
```

### Key takeaway

**Context:** The preferred model was a negative binomial model with number of change efforts as the outcome, including a set of covariates related to structural/contextual elements, student composition, academic performance, and leader attitude. It further incorporated state fixed effects, which allows us to explore within-state differences. Key coefficients were stable across the different model specifications.  

**Finding:** The scope of change efforts underway in districts appears closely linked to district type, leadership tenure, and organizational focus. Districts with long-tenured leaders generally pursue fewer change efforts, whereas districts that prioritize student engagement tend to engage in a larger number of concurrent change efforts. These patterns hold even after accounting for differences in district context and demographics. This suggests that leadership stability and district priorities shape how districts approach change, particularly in terms of pace and breadth.  

I've created one last table to highlight this pattern - the table below provides the predicted number of change efforts based on these characteristics.  

```{r}
# Pull AME predictions
pred_dist <- avg_predictions(
  m5,
  by = "district_type",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = district_type)
pred_tenure <- avg_predictions(
  m5,
  by = "length_tenure",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = length_tenure)
pred_engage <- avg_predictions(
  m5,
  by = "student_engagement",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = student_engagement) %>% 
  mutate(var = ifelse(var == 1, "Selected increase student engagement", "Did not select increase student engagement"))
pred_quality <- avg_predictions(
  m5,
  by = "instructional_quality",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = instructional_quality) %>% 
    mutate(var = ifelse(var == 1, "Selected improve academic achievement", "Did not select improve academic achievement"))
combine <- bind_rows(pred_dist, pred_tenure, pred_engage, pred_quality) %>% 
  select(-df) %>% 
  mutate(estimate = round(estimate, 2))
# Table
datatable(
  combine,
  caption = "Table X. Predicted number of change efforts by district characteristics and reported priorities. 
  Values are model-based average predictions of the expected number of change efforts for each group, 
  holding other factors constant on average.",
  colnames = c("Group", "Predicted number of change efforts"),
  rownames = FALSE
)
```

```{r}
# Clean up
rm(combine, m0, m1, m2, m3, m4, m5, desc, pred_dist, pred_engage, pred_quality, pred_tenure, variance, vars, mod_cc)
```

## Dichotomous

Our dichotomous specification uses original survey responses for Q6, which indicate up to 9 potential areas for change efforts that a district may be taking on, selected from 17 possible responses. We've split them into two categories: fewer change efforts (0-2) and more change efforts (3+).  

I use a logistic regression model to account for the binary outcome, with results presented in exponentiated form similarly to the negative binomial models above, but in odds ratios (rather than as a log of the expected count). In this type of model, a value of 1 indicates no difference in the odds. Values greater than 1 indicate higher odds of reporting high levels of change, while values less than 1 indicate lower odds. For example, a value of 2 indicates that the odds are twice as high, whereas a value of 0.5 indicates that the odds are half as high, holding other factors constant.    

The table below shows the descriptive statistics for this specification of the outcome variable.  

```{r}
desc <- psych::describe(clean$change_group_dich) %>% as_tibble()
variance <- round(var(clean$change_group_dich), 3)
desc %>% 
  mutate(
    variable = "N change efforts",
    variance = variance) %>% 
  select(variable, n, mean, sd, min, max, se, variance) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  DT::datatable()
```

Checking collinearity before modeling:

Checking collinearity:

Again, for the binary model the only meaningful collinearity in the model is between ELA and math achievement; all other predictors exhibit low variance inflation. We could consider creating an index/composite measure if we want to reduce noise and be able to say something about acheivement, but otherwise should just be cautious with interpretation and use them only as controls/not for interpretation. We could consider the same for % low income and % black/latine, but these are within an acceptable level (<3).

```{r}
vif_mod <- lm(
  change_group_dich ~ 
    district_type + c_enrollment + c_enrollment_change + grades_served +
    c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped +
    c_ela_3_8 + c_math_3_8 + c_grad +
    c_breadth_goals + c_instructional_change +
    c_systemic_change + c_will_confidence,
  data = mod
)
vif(vif_mod)
```

The table below presents the modeling results tests the association between the number of efforts a district is working on with each block of covariates one at a time.   

Some noticings from the table:  
-other/traditional twice as likely to be in the high change category  
-higher % black/latine students lowers odds a bit (n.s.)  
-higher % EL/SPED increases odds a bit (n.s.)  
-goal to increase student engagement doubles odds (sig.)  
-tenure is interesting again - less than 1 year nearly doubles the odds, then that goes down until 10+ years halves the odds (n.s.)  

```{r}
# Build models
m0 <- feglm(change_group_dich ~ 1, data = mod, family = binomial())
m1 <- update(m0, . ~ + urbanicity + district_type + c_enrollment + serve_hs)
m2 <- update(m0, . ~ + c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m0, . ~ + c_ela_3_8 + c_math_3_8)
m4 <- update(m0, . ~ + length_tenure + student_engagement + instructional_quality + leadership_influence + personal_motivation + confidence_leading_change)
```

```{r}
stepwise_table(m0, m1, m2, m3, m4)
```

Next, I've integrated the composites into a single model in the table below, again using a stepwise approach. For this model, to be able to compare model fit, I've limited the sample to the 204 districts with complete data on all covariates. The fully specified model had the strongest fit.  

In this version, the associations are even more exaggerated. Notably, other/traditional districts have WAY higher odds of being in the high-change group and the relationship with tenure is still not significant but the coefficients are bigger.  

```{r}
# Ensure sample is the same
vars <- c(
  "change_group_dich",
  "urbanicity", "district_type", "c_enrollment", "serve_hs",
  "c_pct_blacklatine", "c_pct_lowincome", "c_pct_el", "c_pct_sped",
  "c_ela_3_8", "c_math_3_8",
  "length_tenure", "student_engagement", "instructional_quality", "leadership_influence", "personal_motivation", "confidence_leading_change", 
  "state"
)
mod_cc <- mod %>%
  select(all_of(vars)) %>%
  filter(complete.cases(.))
```


```{r}
# Build stepwise models
m0 <- feglm(change_group_dich ~ 1, data = mod_cc, family = binomial())
m1 <- update(m0, . ~ . + urbanicity + district_type + c_enrollment + serve_hs)
m2 <- update(m1, . ~ . + c_pct_blacklatine + c_pct_lowincome + c_pct_el + c_pct_sped)
m3 <- update(m2, . ~ . + c_ela_3_8 + c_math_3_8)
m4 <- update(m3, . ~ . + length_tenure + student_engagement + instructional_quality + leadership_influence + personal_motivation + confidence_leading_change)
```

```{r}
# display table
stepwise_table(m0, m1, m2, m3, m4)
```

And finally, I run the fully specified model once more with state fixed effects to control for variation across context.

```{r}
m5 <- update(m4, . ~ . | state)
```

```{r}
modelsummary(list(
  "Fully specified" = m4,
  "+ State FE" = m5
),
exponentiate = TRUE,
coef_map = c(
  "(Intercept)" = "Intercept",
  "district_typeOther type" = "Other type",
  "district_typeTraditional" = "Traditional",
  "district_typeCharter" = "Charter",
  "c_enrollment" = "Enrollment",
  "urbanicitySuburban" = "Suburban",
  "urbanicityUrban" = "Urban",
  "urbanicityRural" = "Rural",
  "c_pct_blacklatine" = "% Black/Latine",
  "c_pct_lowincome" = "% Low income",
  "c_pct_el" = "% EL",
  "c_pct_sped" = "% SPED",
  "c_ela_3_8" = "% ELA Proficient",
  "c_math_3_8" = "% Math Proficient",
  "confidence_leading_change" = "Confidence in ability to lead change",
  "personal_motivation" = "Personal motivation to lead change",
  "leadership_influence" = "Believes leadership positively influences outcomes now",
   "instructional_quality" = "Goal to improve academic achievement",
   "student_engagement" = "Goal to increase student engagement",
   "length_tenureLess than 1 year" = "Tenure less than 1 year",
   "length_tenure4-6 years" = "Tenure 4-6 years",
   "length_tenure7–10 years" = "Tenure 7–10 years",
   "length_tenureMore than 10 years" = "Tenure more than 10 years"
),
coef_omit = "theta",
gof_omit = "BIC|Log\\.Lik\\.|R2|Adj\\.|Pseudo|RMSE",
stars = TRUE,
estimate = "{estimate}{stars} ({std.error})",
statistic = NULL,
fmt = 2)
```

### Key takeaway

**Context:** The preferred model was a logistic regression model with a dichotomous measure of number of change efforts as the outcome (1 = 3+ change efforts), including a set of covariates related to structural/contextual elements, student composition, academic performance, and leader attitude. It further incorporated state fixed effects (FE), which allows us to explore within-state differences. Key coefficients were roughly stable across model specifications with a little shrinking once we add state FE, which is to be expected.  

**Finding:** Districts that are undertaking a larger number of simultaneous change efforts differ systematically in district type, leader tenure, and organizational priorities. Districts led by long-tenured leaders are less likely to pursue three or more concurrent change efforts, while districts that prioritize student engagement and academic improvement are more likely to report higher levels of change activity. These patterns hold even after accounting for district context, demographics, and state-level differences.  

We have one weird estimate for the belief that leadership currently influences outcomes. This produced an estimated association of zero in the state fixed effects model (and the one without FE). This reflects extreme imbalance in the underlying measure, with nearly all respondents expressing agreement, and only 2 who disagreed. As a result, this estimate should be interpreted as lacking sufficient variation to reliably estimate a relationship.  

I've created one last table to highlight this pattern - the table below provides the predicted probability that a district would fall into our "high change effort" group based on these characteristics.  

```{r}
# Pull AME predictions
pred_dist <- avg_predictions(
  m5,
  by = "district_type",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = district_type)
pred_tenure <- avg_predictions(
  m5,
  by = "length_tenure",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = length_tenure)
pred_engage <- avg_predictions(
  m5,
  by = "student_engagement",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = student_engagement) %>% 
  mutate(var = ifelse(var == 1, "Selected increase student engagement", "Did not select increase student engagement"))
pred_quality <- avg_predictions(
  m5,
  by = "instructional_quality",
  type = "response",
  vcov = FALSE
) %>% 
  rename(var = instructional_quality) %>% 
    mutate(var = ifelse(var == 1, "Selected improve academic achievement", "Did not select improve academic achievement"))
combine <- bind_rows(pred_dist, pred_tenure, pred_engage, pred_quality) %>% 
  select(-df) %>% 
  mutate(estimate = round(estimate, 2))
# Table
datatable(
  combine,
  caption = "Table X. Predicted probability of high change efforts by district characteristics and reported priorities. 
  Values are model-based average predictions of the expected number of change efforts for each group, 
  holding other factors constant on average.",
  colnames = c("Group", "Predicted probability"),
  rownames = FALSE
)
```

```{r}
# Clean up
rm(combine, desc, m0, m1, m2, m3, m4, m5, mod_cc, pred_dist, pred_engage, pred_quality, pred_tenure, vif_mod, variance)
```

